### 2015-04-04 p1_505/pres/pres.tex
first (combining all the old repos and adding project 3)

\documentclass[mathserif,18pt,xcolor=table]{beamer}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{ulem}
\usepackage{slashed}
\usepackage{graphicx}
\usepackage{textpos}
\usepackage{listings}
\usepackage{epsfig}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{fixltx2e}
\definecolor{dukeblue}{RGB}{0,0,156}
\definecolor{dukedarkblue}{RGB}{0,26,87}
\definecolor{dukeblack}{RGB}{79,79,79}
\definecolor{dukegray}{RGB}{79,79,79}
\definecolor{dukesecbrown}{RGB}{217,200,158}
\definecolor{dukesecblue}{RGB}{127,169,174}
\mode<presentation> {
  \usetheme{Boadilla}  
  \setbeamercovered{invisible}
  \setbeamertemplate{navigation symbols}{}  
  \setbeamertemplate{frametitle}[default][center]
  \setbeamerfont{title}{series=\bfseries,parent=structure}
  \setbeamerfont{frametitle}{series=\bfseries,parent=structure}
  \setbeamerfont{subtitle}{size=\scriptsize,series=\bfseries,parent=structure}
  \setbeamerfont{author}{size=\scriptsize,parent=structure}
  \setbeamerfont{institute}{size=\small,series=\bfseries,parent=structure}
  \setbeamerfont{date}{size=\scriptsize,parent=structure}
  \setbeamerfont{footline}{size=\tiny,parent=structure}
  \setbeamercolor{normal text}{bg=white,fg=dukeblack}
  \setbeamercolor{structure}{fg=dukeblue}
  \setbeamercolor{alerted text}{fg=red!85!black}
  \setbeamercolor{item projected}{use=item,fg=black,bg=item.fg!35}
  \setbeamercolor*{palette primary}{use=structure,fg=white, bg=dukeblue}
  \setbeamercolor*{palette secondary}{use=structure,bg=dukedarkblue,fg=white}
  \setbeamercolor*{framesubtitle}{fg=dukegray}
  \setbeamercolor*{block title}{parent=structure,fg=black,bg=dukeblue}
  \setbeamercolor*{block body}{fg=black,bg=dukeblack!10}
  \setbeamercolor*{block title alerted}{parent=alerted text,bg=black!15}
  \setbeamercolor*{block title example}{parent=example text,bg=black!15}
}

\makeatletter
\setbeamertemplate{footline}{
  \leavevmode
  \hbox{%
    \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
      \usebeamerfont{author in head/foot}\insertshortauthor%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
      \usebeamerfont{title in head/foot}\insertshorttitle%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot}%
      \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}%
      \insertframenumber{} / \inserttotalframenumber\hspace*{2ex}%
  \end{beamercolorbox}}%
  \vskip0pt%
}
\makeatother

\AtBeginSection{\frame{\sectionpage}}

\defbeamertemplate{section page}{mine}[1][]{%
  \begin{centering}
    {\usebeamerfont{section name}\usebeamercolor[fg]{section name}#1}
    \vskip1em\par
    \begin{beamercolorbox}[sep=12pt,center]{part title}
      \usebeamerfont{section title}\insertsection\par
    \end{beamercolorbox}
  \end{centering}
}

\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{amsmath}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\pgfdeclareimage[height=1.7cm]{atlaslogo}{logos/atlas_logo.pdf}
\pgfdeclareimage[height=1.7cm]{dukelogo}{logos/dukelogo.pdf}
\title[PHY 505 Project 1]{An Exercise in Geant4}
\author[DD, ME, JR, PZ]{{\small Douglas Davis, Matthew Epland, Justin Raybern, Pingchuan Zhao}}
\institute{\it{Duke University} \\ \mbox{} \\ \mbox{} \\ \pgfuseimage{dukelogo}}
\date[\today]{5 February 2014}
\begin{document}
\defverbatim[colored]\lstI{
  {\scriptsize
    \begin{lstlisting}[language=C++,basicstyle=\ttfamily,keywordstyle=\color{blue}]
      G4RunManager* runMan = new G4RunManager;

      runMan->SetUserInitialization(new MyDetectorConstruction);
      runMan->SetUserInitialization(new MyPhysicsList);
      runMan->SetUserInitialization(new MyActionInitialization);
      runMan->Initialize();

      G4UImanager* UI = G4UImanager::GetUIpointer();
      UI->ApplyCommand(``...'');

      runMan->BeamOn(numberOfEvents);
  \end{lstlisting}}
}

\beamertemplateballitem
\frame{\titlepage}
\addtobeamertemplate{frametitle}{}{}

\begin{frame}
  \frametitle{Outline}
  \begin{itemize}
  \item Introduction to Geant4
  \item Minimum Geant4 Simulation Requirements
  \item Modified Example
  \item Simple Results
  \item Live Demo
  \item Summary
  \end{itemize}
  \centerline{
    \includegraphics[width=.3\linewidth]{pics/g4.pdf}}
\end{frame}

\begin{frame}
  \frametitle{Introduction to Geant4}
  \begin{itemize}
  \item Geant4 is a ``a toolkit for the simulation of the passage of particles through matter.''
  \item Geant4 has broad applications -- high energy and nuclear physics, accelerator physics, and medical and space sciences.
  \item Completely written in C++; successor to GEANT3 (part of \texttt{CERNLIB} written in Fortran).
  \item In development since the 90's.
  \item Visualization, multithreading, and analysis tools exist.
  \item Requires only a C++ compiler and CMake.
    \begin{itemize}
    \item Optional features (such as visualization drivers) require external headers and libraries (e.g. as OpenGL X11)
    \end{itemize}
  \item Many physics lists exist which are tailored for certain processes and certain energy regions.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Minimum Geant4 Simulation Requirements}
             {\footnotesize
               \begin{itemize}
               \item The \texttt{main()} function in a Geant4 simulation requires a minimum of three classes from the Geant4 Libraries: \texttt{G4DetectorConstruction}, \texttt{G4PhysicsList}, \texttt{G4ActionInitialization}.
               \item These objects are given to \texttt{G4RunManager}.
               \item A pointer to the UI manager can be created to supply user commands to the Geant4 Kernel (for example \texttt{/trackings/verbose 1} to have the program print all tracking information)
               \item \texttt{G4RunManager::BeamOn} begins the simulation process.
             \end{itemize}}
             \lstI
\end{frame}

\begin{frame}
  \frametitle{\texttt{DetectorConstruction}}
  \begin{itemize}
  \item \texttt{G4WorldVolume} -- The volume in which all other volume instances must live.
  \item \texttt{G4LogicalVolume} -- A shape of some material with physical characteristics. The blueprint of a logical volume class is where volumes can be made sensitive (we do this). Logical volumes must be defined with a \texttt{G4Material}.
  \item \texttt{G4PhysicalVolume} -- Instances of Logical Volumes that live inside the world volumes, multiple physical volumes can be made from a single logical volume.
  \end{itemize}
\centerline{\includegraphics[width=.7\linewidth]{pics/volumes.png}}
\end{frame}

\begin{frame}
  \frametitle{\texttt{PhysicsLists} and \texttt{ActionInitialization/PrimaryGeneratorAction}}
  {\footnotesize
  \begin{itemize}
  \item Physics lists are provided by the Geant4 collaboration.
  \item The collaboration provides descriptions of each physics list which details what kind of physical processes and energy ranges are best for the different lists -- the user then chooses accordingly.
  \item \texttt{G4UserPhysicsList} allows the user to define different particles, processes, and also range cuts of parameters
  \item The user defines the intial state of the system in the \texttt{UserActionInitialization/UserPrimaryGeneratorAction}. What particles should initialize the event, the definition of electric and magnetic fields, etc.
  \item These classes are all derived classes with virtual methods that must be defined by the user for their specific needs!
  \end{itemize}}
\end{frame}

\begin{frame}
  \frametitle{\texttt{RunAction} and \texttt{EventAction}\\(Examples of Optional Classes)}
  \begin{itemize}
  \item A Geant4 run consists of events that all share the same detector, physics, and initial system.
  \item \texttt{RunAction} allows the user to define what occurs at the beginning of a run and at the end of a run using the functions \texttt{RunAction::BeginOfRunAction} and \texttt{RunAction::EndOfRunAction}. The RunAction is made to always call these functions. The user can also create his/her own functions to be called during the run.
  \item \texttt{EventAction} is very similar to RunAction, but of course is initialized $N$ times. 
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Our Modified Example}
  {\footnotesize
  \begin{itemize}
  \item Our simple simulation is based off of \texttt{exampleB4c} from the Geant4 examples.
  \item The example fires a 50 MeV electron into a calorimeter made of a variable number of layers of lead and liquid argon.
  \item B4c yields the total energy deposited and track length in all layers of lead and the total energy deposited in all layers liquid argon combined (no independent layer information). (Using the \texttt{G4Hit} class)
  \item We modified the example in the following way:
    \begin{itemize}
    \item 8 independent ``channels'' (\texttt{G4SensitiveDetectors}) of variable material type (lead, polypropylene, or water) were used in our simple exercise.
    \item Store information on a per layer basis in a \texttt{ROOT} ntuple for each event using a pointer to \texttt{G4RunAction} in \texttt{G4EventAction}.
    \item Particle type and energy modifications. (We used a 2 GeV proton, neutron, muon, and electron)
    \item Two different physics lists were employed. (\texttt{FTFP\_BERT}, \texttt{QGSP\_BERT}).
    \item All 24 scenarios were run with 10,000 events each.
    \end{itemize}
  \end{itemize}}
\end{frame}

\begin{frame}
  \frametitle{Our Modified Example cont'd}
  {\footnotesize
  \begin{itemize}
    \item Distributions of E\textsubscript{deposited} and average track length per layer were created from each layers sensitive detector readout.
    \item Both physics lists produced identical results for the energies, materials, and particles considered. Therefore, we'll discuss 12 systems in the simple analysis of the upcoming slides.
    \item The following plots were all created from simulations using the \texttt{FTFP\_BERT} physics list.
  \end{itemize}}
\end{frame}

% electron
\begin{frame}
  \frametitle{Electron Interactions}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/e-Pb.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/electron_pb_edep.pdf}}
  \put(18.5, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/e-PP.png}}
  \put(18.5, -60){\includegraphics[width=.48\textwidth]{../report/plots/electron_pp_edep.pdf}}
  \end{picture}
\end{frame}

\begin{frame}
  \frametitle{Electron Interactions cont'd}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/e-H2O.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/electron_h2o_edep.pdf}}
  \put(18.5, -30){\includegraphics[width=.48\textwidth]{../report/plots/TL_electron.pdf}}
  \end{picture}
\end{frame}

% muon
\begin{frame}
  \frametitle{Muon Interactions}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/mu-Pb.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/muon_pb_edep.pdf}}
  \put(18.5, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/mu-PP.png}}
  \put(18.5, -60){\includegraphics[width=.48\textwidth]{../report/plots/muon_pp_edep.pdf}}
  \end{picture}
\end{frame}

\begin{frame}
  \frametitle{Muon Interactions cont'd}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/mu-H2O.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/muon_h2o_edep.pdf}}
  \put(18.5, -30){\includegraphics[width=.48\textwidth]{../report/plots/TL_muon.pdf}}
  \end{picture}
\end{frame}

% proton
\begin{frame}
  \frametitle{Proton Interactions}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/P-Pb.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/P_pb_edep.pdf}}
  \put(18.5, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/P-PP.png}}
  \put(18.5, -60){\includegraphics[width=.48\textwidth]{../report/plots/P_pp_edep.pdf}}
  \end{picture}
\end{frame}

\begin{frame}
  \frametitle{Proton Interactions cont'd}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/P-H2O.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/P_h2o_edep.pdf}}
  \put(18.5, -30){\includegraphics[width=.48\textwidth]{../report/plots/TL_proton.pdf}}
  \end{picture}
\end{frame}

% neutron
\begin{frame}
  \frametitle{Neutron Interactions}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/N-Pb.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/N_pb_edep.pdf}}
  \put(18.5, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/N-PP.png}}
  \put(18.5, -60){\includegraphics[width=.48\textwidth]{../report/plots/N_pp_edep.pdf}}
  \end{picture}
\end{frame}

\begin{frame}
  \frametitle{Neutron Interactions cont'd}
  \begin{picture}(320,250)(-160,-125)
  \put(-160, 60){\includegraphics[width=.48\textwidth, trim = 0mm 75mm 0mm 75mm, clip]{../report/pics/N-H2O.png}}
  \put(-160, -60){\includegraphics[width=.48\textwidth]{../report/plots/N_h2o_edep.pdf}}
  \put(18.5, -30){\includegraphics[width=.48\textwidth]{../report/plots/TL_neutron.pdf}}
  \end{picture}
\end{frame}

\setbeamertemplate{section page}[mine]
\section{Live Demo}

\begin{frame}
  \frametitle{Summary}
  {\footnotesize
  \begin{itemize}
    \item Geant4 is a powerful simulation tool written in C++ for many research areas (notably particle and nuclear, medical, and space physics).
    \item We discussed some of the required and major features in a Geant4 simulation program
    \item A simple simulation program was shown which was developed based on a modifed Geant4 collaboration provided example
    \item We observed interactions that are expected based on our knowledge of physics.
    \item Code for entire project: \url{http://github.com/dougphy/p1_505}
  \end{itemize}}
\end{frame}

\end{document}

### 2015-04-04 p1_505/report/paper1.tex
first (combining all the old repos and adding project 3)

\documentclass[aps,prd,twocolumn,nofootinbib]{revtex4-1}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{epsfig}
\usepackage{listings}
\usepackage[hidelinks,hyperfootnotes=false,bookmarks=false]{hyperref}
\begin{document}
\title{An Exercise in Geant4}
\author{Douglas Davis}
\author{Matthew Epland}
\author{Justin Raybern}
\author{Pingchuan Zhao}
\affiliation{Department of Physics, Duke University, Durham, NC 27707, USA}
\date{\today}
\begin{abstract}
In this report we introduce the essential components of a Geant4 simulation program and describe a simple Geant4 simulation that was developed using a standard Geant4 collaboration suppled example. We investigate the interaction of four different particles (electron, muon, proton, and neutron) in a block of three different materials (lead, polypropylene, and water). The material blocks are separated into eight individual segments (which can be imagined as eight different channels in a real detector). Event displays for the interactions are shown and qualitatively discussed. Quantitative analysis is done by investigating energy deposition and track lengths of the particles in each of the eight channels. Distributions of energy deposition per layer and track length per layer are shown and discussed.
\end{abstract}\maketitle
\section{Motivation and Description}
\label{sec1}
Geant4~\cite{geant} is the current standard in simulating nuclear and particle interactions. It is used extensively by researchers of many fields in physics –- including nuclear and high energy physics as well as astronomy, medical and space physics. The flexibility that the user has to control the environment and physical processes makes it a valuable tool for simulating nearly any experiment involving particle interactions. The software package has been in development since the 1990's, and is the successor to the \texttt{CERNLIB} package \texttt{GEANT3}, written in Fortran. We use the most recent version of Geant4 (version 10.1) which only requires a \texttt{C++} compiler and the CMake program to install. Optional features in a Geant4 installation include visualization drivers, multi-threading support, and other more advanced options. These optional features require other external headers and libraries. For our purposes, we use the OpenGL X11 optional feature for visualizing events. In this project we investigated some simple capabilities of Geant4 by propagating various particles through a number of materials. Our project is derived from one of the default Geant4 examples~\cite{G4_tutorials}.

Geant4 is written in \texttt{C++} and makes full use of \texttt{C++}'s object oriented architecture. Each simulation must define any primary particles, the detector materials and construction (i.e. the detector apparatus itself plus the surrounding environment including any magnetic fields, etc.), and the types of physics processes that should be modeled. These definitions are handled in the \texttt{G4ActionInitialization}, \texttt{G4DetectorConstruction}, and \texttt{G4PhysicsList} classes respectively. In addition to these three main classes a few additional classes, such as \texttt{G4SteppingAction}, \texttt{G4RunAction}, and \texttt{G4SensitiveDetector}, are also needed to have a fully fledged simulation that can propagate particles and provide output data from the detecting material. This of course is on top of the code base and libraries provided by the Geant4 instillation itself. Skeleton working examples and tutorial simulations are readily available, so it is usually not necessary to start creating a Geant4 simulation from scratch.

\section{Simulation and Results}
\subsection{Simulation Setup}
In order to gain familiarity with Geant4 we derived a simple simulation from the B4c sampling calorimeter example. The original B4c example consisted of a variable number of layers of a calorimeter made from lead and liquid argon with a 50 MeV electron propagating through. The lead and liquid argon volumes are defined by inheritance from \texttt{G4SensitiveDecector}. This allows for the definition of a specific hit class inheriting from \texttt{G4Hit}, which is used to store the energy and track length in all of the lead and liquid argon. There is no independence between the different layers of the materials. The simulation uses the \texttt{FTFP\_BERT} physics list (FRITIOF description of string excitation and fragmentation with Geant4 Bertini cascade for primary protons, neutrons, pions and Kaons below approximately 10 GeV).

Modifications to the example for our project include changes to the \texttt{PrimaryGeneratorAction}, \texttt{PhysicsList} \texttt{DetectorConstruction}, \texttt{RunAction}, and \texttt{EventAction}. For the generation of particles, we change the energy of the incoming particle, and also vary the particle type. We use electrons, muons, protons, and neutrons. For the detector construction, we create what is essentially a single block of material, with eight independent ``channels.'' When the particle propagates through the detector, it is always going through the same material, but independent layer by layer hit information is created by declaring eight different \texttt{LogicalVolume} objects.\footnote{For further discussion of volumes in Geant4 see Appendix~\ref{sec:AppendixA}} At the beginning of the \texttt{RunAction}\footnote{For further discussion of run and event actions, see Appendix~\ref{sec:AppendixB}} a ROOT~\cite{ROOT} \texttt{TFile} is initialized along with a \texttt{TTree} for data output. Sixteen branches are created for storing the energy deposited and track length for each layer. In the \texttt{EventAction}, \texttt{G4HitsCollection} pointers are declared to retrieve the total energy deposited and track length in each layer. These objects are pulled from the \texttt{G4SensitiveDetector} Manager. A pointer to the \texttt{RunAction} is declared in the \texttt{EventAction} where the function \texttt{RunAction::UpdateTree} has the current event's energy deposition and track length values updated in the \texttt{RunAction}, filling the whole run's \texttt{TTree}. At the end of the number of requested events, the \texttt{RunAction::EndOfRunAction} function writes the ROOT \texttt{TTree} to the \texttt{TFile} and closes the \texttt{TFile}.

Two Geant4 run macros were used. Geant4 macros are a way to send commands to the Geant4 kernel. The default macro initialized the visualization manager, drawing the detector volume in an OpenGL canvas. The simple command \texttt{/run/beamOn N} would initialize the \texttt{RunAction} and have the kernel run $N$ events. For visualizing events, $N=1$ suffices. Figure~\ref{fig:e-geo} shows an event displays for a 2 GeV electron created by executing the program with the default macro and running the \texttt{/run/beamOn 1}. Visualizing events is very CPU heavy; therefore, for generating large data samples a macro which would not initialize the visualization was used. This macro called the kernel to run 10,000 events and at the end of the run exit the Geant4 kernel.

\subsection{Event Visualizations and Qualitative Event Analysis}

As previously described, visualizing events was done using the OpenGL X11 driver in Geant4. We generated event displays for all four particles in all three materials. Figures~\ref{fig:e-geo},~\ref{fig:mu-geo},~\ref{fig:P-geo},~and~\ref{fig:N-geo} show event displays for electron, muon, proton, and neutron events in lead, polypropylene, and water.

Examination of the event displays yields an estimate-like understanding of the event. We can see that passing through lead very much stops the electron, proton, and neutron, while the muon is able to pass through completely. The yellow dots along the track shows that the muon does interact more in the lead than it does in polypropylene and water, but it still passes through the complete detector volume. Examining the electron in water event closely appears to have a candidate Cherenkov radiation cone. One would possibly expect to see a Cherenkov cone in the muon in water event, but that is not present. This may be an issue with the physics list choice; and could be an item of deeper investigation with more time.

\begin{figure*}
  \centering
  \subfloat[Electron events \label{fig:e-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/electron_ed.png}}
  \hfill
  \subfloat[Muon events \label{fig:mu-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/muon_ed.png}}
  \hfill
  \subfloat[Proton events \label{fig:P-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/proton_ed.png}}
  \hfill
  \subfloat[Neutron events \label{fig:N-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/neutron_ed.png}}
  \caption{The detector geometry with a 2 GeV particle events in three different materials. The top displays corresponds to lead, the middle displays to polypropylene, and the bottom displays to water. Red trajectories indicate negatively charged tracks, green neutral tracks, and blue positively charged tracks.}
  \label{fig:mean and std of nets}
\end{figure*}

\subsection{Quantitative Results}

The deposited energy and average track length per layer plots generally reinforce the impressions gained from viewing the visualizations. The Pb absorbs all four types of particles very rapidly while the polypropylene and water do not halt the particles as fast and behave very similarly to each other. This is likely due to both of them containing large numbers of H nuclei, bound to either O or C, for the incoming particles to interact with.

One interesting artifact in the plots can be seen in the average track length per layer plot for the neutron runs. For both polypropylene and water the first bin for 0 track length has an enormous excess of events compared to the rest of the plot. Due to time constraints we were unable to determine the cause of this excess in detail, but due to the picture presented by the deposited energy plots and visualizations it is not likely that these neutrons are really being completely stopped by the polypropylene and water. The spike may be an artifact of the normalization procedure run on this plot.

\begin{figure*}
  \includegraphics[width=.45\textwidth]{plots/electron_pb_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/electron_pp_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/electron_h2o_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/TL_electron.pdf}
  \caption{Electron Plots, Pb $E_{\text{dep}}$ plot effectively has units of GeV not MeV, the $10^3$ is obscured.}
  \label{fig:e-quant}
\end{figure*}

\begin{figure*}
  \includegraphics[width=.45\textwidth]{plots/muon_pb_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/muon_pp_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/muon_h2o_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/TL_muon.pdf}
  \caption{Muon Plots.}
  \label{fig:mu-quant}
\end{figure*}
\begin{figure*}
  \includegraphics[width=.45\textwidth]{plots/P_pb_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/P_pp_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/P_h2o_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/TL_proton.pdf}
  \caption{Proton Plots, Pb $E_{\text{dep}}$ plot effectively has units of GeV not MeV, the $10^3$ is obscured.}
  \label{fig:p-quant}
\end{figure*}

\begin{figure*}
  \includegraphics[width=.45\textwidth]{plots/N_pb_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/N_pp_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/N_h2o_edep.pdf}
  \includegraphics[width=.45\textwidth]{plots/TL_neutron.pdf}
  \caption{Neutron Plots, Pb $E_{\text{dep}}$ plot effectively has units of GeV not MeV, the $10^3$ is obscured.}
  \label{fig:n-quant}
\end{figure*}

\section{Conclusions}
This report introduced the very powerful, \texttt{C++} based  Geant4 software package for the simulation of the passage of particles through matter. We discussed the minimum requirements for a working Geant4 simulation, and also described some optional features of Geant4. We discussed our development of a simple system which was born from a Geant4 collaboration provided example. Some quantitative results were shown and discussed, and some qualitative discussion was presented. Some expected distributions were observed, while event displays showed some expected results as well. The code for this project can be found here:\\ \url{http://github.com/dougphy/p1_505}.

\begin{thebibliography}{9}
\bibitem{geant}
  GEANT4 Collaboration, S. Agostinelli et al., \textit{GEANT4: A Simulation toolkit}, Nucl. Instr. Meth. \textbf{A506} (2003) 250.
\bibitem{G4_tutorials}
  GEANT4 Collaboration, S. Agostinelli et al., \url{http://geant4.web.cern.ch/geant4/UserDocumentation/Doxygen/examples_doc/html/index.html}
\bibitem{ROOT}
  R.~Brun and F.~Rademakers, Nucl.\ Instrum.\ Meth.\ A {\bf 389}, 81 (1997). \texttt{http://root.cern.ch/}
\end{thebibliography}

\begin{appendix}

\section{Volumes in Detector Construction}\label{sec:AppendixA}
There are three definitions of volumes that need to be understood in detector construction. The ``world volume'' is the volume in which everything exists. The user defines the shape and properties of the world volume and then everything of interest, including any physical objects or shot particles, must be contained inside the world volume. If a particle track takes it outside the world volume, it is no longer considered part of the simulation. A ``logical volume'' has a physical shape and characteristics such as material and dimensions. This might be a block of NaI or a sphere of air. A ``physical volume'' is an instance of a logical volume. This volume is given coordinates within the world volume. The user may make one or more copies of a logical volume within the world volume (meaning one or more physical volumes). These are the actual blocks of matter in which the simulation will occur.
\section{Run and Event Actions}\label{sec:AppendixB}
The \texttt{RunAction} class in Geant4 gives the user the ability to control what happens at the beginning of an event and at the end of an event. Other member functions can be added to the class to give the user more control of what happens during a run. What is defined in the \texttt{RunAction} cannot be changed on an event by event basis. The \texttt{EventAction} class is similar to \texttt{RunAction} but of course for an event. The user has control of what happens at the beginning of an event and at the end -- and also can add member functions for specific tasks to run during the events. For example, the user can define control statements so different things can happen for a given number of particles, or track length, or amount of energy deposited, etc. in an event.

\end{appendix}

\end{document}

### 2015-04-04 p2_505/paper/paper2.tex
first (combining all the old repos and adding project 3)

\documentclass[aps,prd,twocolumn,nofootinbib]{revtex4-1}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{epsfig}
\usepackage{listings}
\usepackage[hidelinks,hyperfootnotes=false,bookmarks=false]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{verbatim}

%\usepackage[section]{placeins}
\usepackage{float}

\DeclareRobustCommand{\su3flavor}{$\text{SU(3)}_{\text{flavor}}$}
\DeclareRobustCommand{\qqbar}{$q\overline{q}$}

\begin{document}

\title{Mesons Beyond \su3flavor \qqbar}
\author{Douglas Davis}
\author{Matthew Epland}
\author{Justin Raybern}
\author{Pingchuan Zhao}
\affiliation{Department of Physics, Duke University, Durham, NC 27707, USA}
\date{\today}
\begin{abstract}
%\todo[inline]{Abstract goes here!}
Mesons beyond \qqbar~are important for refining our understanding of QCD. Both hybrid and exotic mesons are discussed along with the theoretical motivation of, and experimental evidence for, each. The QCD hybrids would provide an excellent test of lattice QCD predictions. Glueballs and tetraquarks are exotic mesons allowed by QCD and there is experimental evidence of the latter and several experimental candidates for the former. There is much work to do but the results seem to point to an exciting future.
\end{abstract}\maketitle

\section{Theoretical Motivations}
The particles we call mesons are usually introduced as consisting of a single \su3flavor quark-antiquark pair, \qqbar. This description is somewhat simplistic however as the formal definition of a meson is any hadron, a particle made up of quarks, that has baryon number $B=0$.
\begin{align}
\label{eq:1}
B = \frac{1}{3}(n_{q}-n_{\overline{q}})
\end{align}

The $B=0$ definition allows us to group a larger number of theorized particles together under the name mesons; those that can not be fully described as \su3flavor \qqbar~states are typically called exotic mesons. In this paper we will be discussing three of the main types of exotic mesons: QCD Hybrids, Glueballs, and Tetraquarks.

\subsection{QCD Hybrids}
The usual \su3flavor \qqbar~mesons can be well described as being colorless bound states of $q_{i}\overline{q}_{j}$, $q_{i}$ with some color and $\overline{q}_{j}$ with the corresponding anticolor, and can be grouped into isospin triplets based on the constituent quark's flavors. Each meson can further be characterized by it's quantum numbers, namely it's isospin $I$, total angular momentum $J$, and parity $P$. Additionally, the neutrally charge state of each isospin triplet is an eigenstate of the charge conjugation operator $\hat{C}$ with eigenvalue $C$. $P$ and $C$ are functions (\ref{eq:2}) of the constituent quark's relative orbital angular momentum $L$ and their total intrinsic spin $S$ \cite{ketzer}.

\begin{align}
\label{eq:2}
P = (-1)^{L+1},~C = (-1)^{L+S}
\end{align}

As quarks are spin--$\frac{1}{2}$ particles, the total spin can only take values of $S = 0\text{ or }1$, and as usual $L$ and $J$ can take values $L = 0, 1, 2, 3, \ldots$; $J = L+S, L+S-1, \ldots, \left| L-S \right|$. All together, this gives us allowed quantum numbers $J^{PC} = 0^{-+}, 0^{++}, 1^{--}, 1^{+-}, 1^{++}, 2^{--}, 2^{-+}, 2^{++},\ldots$ for the neutral \su3flavor \qqbar~mesons \cite{ketzer}.

The quantum numbers that are not allowed for \su3flavor mesons, such as $J^{PC} = 0^{--}, 0^{+-}, 1^{-+}, 2^{+-},\ldots$ are commonly called exotic quantum numbers and are allowed when using Quantum Chromodynamics (QCD) to model the binding of the quarks in the meson. QCD can produce a wider variety of \qqbar~mesons, including those with exotic quantum numbers, because it takes into account the gluon field mediating the strong force between the two quarks. The gluon field can have quantum numbers of it's own that combine with the quark's in the mesons overall quantum numbers, thereby creating the possibility of mesons with exotic quantum numbers.

In addition to the effect they have on a meson's potential quantum numbers, gluons carry color which allows for meson states to exist where the gluon field binding the quarks is not colorless as is implicitly assumed in the \su3flavor model. Such states are called hybrids. Mesons with multiple \qqbar~pairs are also considered QCD hybrids provided that the valence \qqbar's net color is still canceled by a colored gluon field \cite{ketzer}.

As hybrid mesons are predicted by QCD, but not the \su3flavor model, they are an important area of study in order to experimentally verify the predictions of QCD. Hybrid mesons with non-exotic quantum numbers are experimentally hard to detect as they will only be apparent as an excess of states on top of what is predicted by \su3flavor. Exotic hybrid mesons are therefore an area of great experimental interest as it is relatively easy to measure a mesons $J^{PC}$ quantum numbers.

In order to use quantum chromodynamics to predict the properties of hybrid mesons non-perturbative calculations must be performed. This is by and large done with the lattice method, i.e. lattice QCD. While lattice QCD in theory is able to predict the hybrid mesons properties from first principles, in practice the calculations produce results that have large uncertainties and depend on the exact methods used. Experimentally discovering exotic hybrid meson states, and precisely measuring their properties, is therefore an important test of QCD as well as the lattice methods used to derive it's results.

\subsection{Glueballs}
%\todo[inline]{Justin}
A potential non--\qqbar~meson candidate is made from a collection of gluons bound to one another, or a glueball, in the absence of quarks. The key is that gluons carry color charge and feel the strong force just as quarks do \cite{j1}. Detecting glueballs, however, has proven to be incredibly challenging. They exhibit significant mixing with known mesons. Additionally, the masses of the lightest predicted glueballs, those with $J^{PC} = 0^{++} \text{and}~2^{++}$, are around the same mass as known standard mesons. There are several resonances that are potentially linked to glueballs but the evidence remains inconclusive. Many masses for glueballs can derived from theory using a variety of techniques with the QCD sum rule being one of the more popular approaches \cite{j2}. From QCD we would expect two gluon states to have $J \neq 1$, $C = +1$, $P = \pm1$. Three gluon states can have $C = \pm1$ \cite{j3}.

\subsection{Tetraquarks}
%\todo[inline]{Justin}
Another exotic meson category is a tetraquark, or 4 quark meson. The $qq\overline{q}\overline{q}$~tetraquark is allowed by QCD, and unlike the glueball, has been experimentally detected with the $Z(4430)$ \cite{lhcb}. This should not be thought of as two standard mesons bound to one another (though that is a likely way the tetraquark would decay), but as two quarks and two antiquarks bound as one. In addition to tetraquarks, it has been theorized that there could exist pentaquarks ($qqqq\overline{q}$) and even higher numbers of quark groups.

\section{Experimental Results to Date}

\subsection{The $Z(4430)$ Tetraquark}
In 2007, the Belle Collaboration (a $B$ factory experiment using the KEKB $e^+e^-$ linear collider at $\sqrt{s} = 10.58$ GeV, the $\Upsilon(4S)$ mass.) observed a distinct peak in the invariant mass distribution of $\pi^{\pm}\psi^{\prime}$ samples near 4430~MeV~\cite{belle}. The resonant peak was found in $B\rightarrow K \pi^{\pm}\psi^{\prime}$ decays. Figure~\ref{fig:z4430diag} shows the decay chain of such an interaction.

\begin{figure}[H]
  \centerline{\psfig{file=../figures/Z4430_diagram.pdf, width=0.95\linewidth}}
  \caption{The creation and decay chain of the $Z(4430)$ tetraquark meson.}
  \label{fig:z4430diag}
\end{figure}

Due to the charge of the observed resonance it was impossible for the candidate particle to simply be an exited charmonium state ($c\overline{c}$). Therefore, the observed resonance was a suspected tetraquark state (with quark content $c\overline{c}u\overline{d}$~or $c\overline{c}d\overline{u}$). The $J^P$ of the particle was not measured by Belle, and therefore other hypothesis different from the tetraquark could be made. Another possible state was the $D$ ``molecule.'' A superposition of $D$ meson states forming an isospin multiplet~\cite{z4430_dmolecule}:
\begin{align}
Z^+ &= \frac{1}{\sqrt{2}}\left(D^{*+}\overline{D}^0_1-D^+_1\overline{D}^{*0}\right), \\
Z^0 &= \frac{1}{2}\left(D_1^0\overline{D}^{*0} - D^{*0}\overline{D}_1^0 + D_1^+D^{*-}-D^{*+}D^-_1\right),\\
Z^- &= \frac{1}{\sqrt{2}}\left(D^0_1D^{*-} - D^{*0}D_1^-\right).
\end{align}
Another $B$ factory experiment, BABAR at SLAC, was unable to find the same resonant behavior. Therefore, the official tetraquark discovery was not confirmed by the $B$ factory experiments.

In 2014, the LHCb collaboration at the LHC confirmed the existence of the tetraquark state $Z(4430)^-$ with quark content $c\overline{c}d\overline{u}$~\cite{lhcb}. LHCb confirmed the state at 13.9$\sigma$. The spin-parity of the state was measured to be $J^P = 1^+$, eliminating the possible $D$ molecule and confirming the tetraquark state. The mass and width measured by LHCb are:
\begin{align}
m_{Z^{-}} &= 4475 \pm 7^{+15}_{-25}\text{ MeV},\\
\Gamma_{Z^{-}} &= 172 \pm 13^{+37}_{-34}\text{ MeV}.
\end{align}
Figures~\ref{fig:z4430lhcb1} and ~\ref{fig:z4430lhcb2} show the resonant peak and also the likelihood measurement of for the spin-parity.

\begin{comment}
\begin{figure*}[h!]
  \centerline{\psfig{file=../figures/Z4430_res.pdf, width=.45\linewidth}
  \psfig{file=../figures/Z4430_jp.pdf,width=.45\linewidth}}
  \caption{LHCb plots}
  \label{fig:z4430lhcb}
\end{figure*}
\end{comment}

\begin{figure}[H]
  \centerline{\psfig{file=../figures/Z4430_res.pdf, width=.95\linewidth}}
  \caption{The LHCb measurement of the invariant mass of the $\pi\psi^{\prime}$ system. The peak corresponds to $m=4475$ MeV.}
  \label{fig:z4430lhcb1}
\end{figure}

 \begin{figure}[H]
  \centerline{\psfig{file=../figures/Z4430_jp.pdf,width=.95\linewidth}}
  \caption{The LHCb measurement of the spin-parity of the $Z(4430)$.}
  \label{fig:z4430lhcb2}
 \end{figure}

\subsection{$\pi(1400)$ and $\pi(1600)$}
%\todo[inline]{Ping}

Both the Lattice QCD model and the QCD string (QCD flux tube) models predict an exotic hybrid meson with $J^{PC} = 1^{-+}$ with a mass of around 1.9$\pm$0.2 GeV. However, the bag model predicts the exotic $1^{-+}$ at a lower mass of about 1.4 GeV. This exotic $1^{-+}$ meson decaying to $\eta \pi^{-}$, is called $\pi_{1}(1400)$. It is reported in the reaction $\pi^{-}p \rightarrow \eta \pi^{-}p$ at 18.3 GeV by the E852 collaboration using the Multi-Particle Spectrometer (MPS) at the AGS. $\eta$ is detected through $\gamma \gamma$ decay. Evidence of $\pi_{1}(1400)$ is also found in $\pi^{-}p \rightarrow \eta \pi^{0} n$ at 18.3 GeV at E852.

\subsubsection{$p\overline{p}$~Annihilation}

The crystal barrel collaboration at LEAR also searched for a $1^{−+}$ resonance in the $\eta \pi$ P-wave\footnote{Hybrid mesons are expected to predominantly decay into pairs of S and P-wave mesons. The S-wave, angular momentum quantum number $l=0$, corresponds to a classical head on collision. P, $l=1$ and D-waves, $l=2$ have a higher angular momentum between the initial and final states.} in low energy $p\overline{p}$~annihilation into $\pi\pi\eta$. The $\pi_{1}(1400)$ signal is weak in peripheral reactions and contains lots of backgrounds which mimic other 1400 resonances such as the $a_{2}(1320)$ meson; the signal is stronger in $p\overline{p}$~annihilation. Also, a similar $\pi_{1}(1600)$ is also reported both from many peripheral experiments and $p\overline{p}$~annihilation. A confirmation that they are the particle predicted by the flux tube and lattice QCD model requires further theoretical and experimental work.

%\subsection{Implications for Lattice QCD}
%\todo[inline]{??, we need some references on this!!!}

\section{Conclusions}
%\todo[inline]{Justin, Includes short discussion of exotic mesons Implications for Lattice QCD, could just be a rehash of what is said in the hybrids section}
The study of exotic mesons will contribute a great deal to our understanding of QCD. In particular, work on hybrid mesons can eliminate poor lattice QCD models in favor of those that make better experimental predictions. Glueballs and tetraquarks remain elusive, though success in verifying the Z(4430) gives hope that others may soon follow. The field could be a fruitful area of research for some time.

\newpage

\begin{thebibliography}{9}
%\bibitem{Barnes}
%  T. Barnes, \textit{Exotic Mesons, Theory and Experiment}, arXiv:hep-ph/0007296, \url{http://arxiv.org/abs/hep-ph/0007296}
\bibitem{ketzer}
  Bernhard Ketzer, \textit{Hybrid Mesons}, Proceedings of Science, Sixth International Conference on Quarks and Nuclear Physics, PoS(QNP2012)025, \url{http://pos.sissa.it/archive/conferences/157/025/QNP2012_025.pdf} %\url{http://arxiv.org/abs/1208.5125}
%\bibitem{Olsen}
%  Stephen Lars Olsen, \textit{QCD exotics}, Hyperfine Interactions, \textbf{229}, \url{http://dx.doi.org/10.1007/s10751-014-1061-4}
\bibitem{belle}
  S.~K.~Choi~\emph{et. al.}, \emph{Observation of a Resonancelike Structure in the} $\pi^{+-}\psi^{\prime}$ \emph{Mass Distribution in Exclusive} $B\rightarrow K \pi^{+-}\psi^{\prime}$ \emph{Decays}. Phys. Rev. Lett. \textbf{100}, 142001 8 April 2008.
\bibitem{z4430_dmolecule}
  E.~Braaten~and~M.~Lu, \emph{Line shapes of the Z(4430)} Phys. Rev. D \textbf{79} 051503(R) 11 March 2009
\bibitem{lhcb} % {j4}
  LHCb Collaboration, R. Aaij \textit{et al.}, \textit{Observation of the Resonant Character of the }$Z(4430{)}^{-}$\textit{ State}, Phys. Rev. Lett., \textbf{112}, 22 (2014). \url{http://dx.doi.org/10.1103/PhysRevLett.112.222002}%\url{http://link.aps.org/doi/10.1103/PhysRevLett.112.222002}

\bibitem{j1}
 K.A. Olive \textit{et al.} (Particle Data Group), \textit{Non--\qqbar~Mesons} (Revised March 2006 by C. Amsler), Chin. Phys. C, 38, 090001 (2014), \url{http://pdg.lbl.gov/2006/reviews/nonqqbar_mxxx050.pdf}
\bibitem{j2}
 Vincent Mathieu, Nikolai Kochelev, and Vicente Vento, \emph{The Physics of Glueballs}, Int. J. Mod. Phys. E \textbf{18}, 1 (2009), \url{http://www.worldscientific.com/doi/abs/10.1142/S0218301309012124}
\bibitem{j3}
 Wolfgang Ochs, \emph{The status of glueballs}, J. Phys. G: Nucl. Part. Phys. \textbf{40}, 043001 (2013), \url{http://dx.doi.org/10.1088/0954-3899/40/4/043001}

\end{thebibliography}

\end{document} %%% end of doc %%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simulation and Results}
\subsection{Simulation Setup}

Modifications to the example for our project include changes to the \texttt{PrimaryGeneratorAction}, \texttt{PhysicsList} \texttt{DetectorConstruction}, \texttt{RunAction}, and \texttt{EventAction}. For the generation of particles, we change the energy of the incoming particle, and also vary the particle type. We use electrons, muons, protons, and neutrons. For the detector construction, we create what is essentially a single block of material, with eight independent ``channels.'' When the particle propagates through the detector, it is always going through the same material, but independent layer by layer hit information is created by declaring eight different \texttt{LogicalVolume} objects.\footnote{For further discussion of volumes in Geant4 see Appendix~\ref{sec:AppendixA}} At the beginning of the \texttt{RunAction}\footnote{For further discussion of run and event actions, see Appendix~\ref{sec:AppendixB}} a ROOT~\cite{ROOT} \texttt{TFile} is initialized along with a \texttt{TTree} for data output. Sixteen branches are created for storing the energy deposited and track length for each layer. In the \texttt{EventAction}, \texttt{G4HitsCollection} pointers are declared to retrieve the total energy deposited and track length in each layer. These objects are pulled from the \texttt{G4SensitiveDetector} Manager. A pointer to the \texttt{RunAction} is declared in the \texttt{EventAction} where the function \texttt{RunAction::UpdateTree} has the current event's energy deposition and track length values updated in the \texttt{RunAction}, filling the whole run's \texttt{TTree}. At the end of the number of requested events, the \texttt{RunAction::EndOfRunAction} function writes the ROOT \texttt{TTree} to the \texttt{TFile} and closes the \texttt{TFile}.

\subsection{Event Visualizations and Qualitative Event Analysis}

\begin{figure*}
  \centering
  \subfloat[Electron events \label{fig:e-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/electron_ed.png}}
  \hfill
  \subfloat[Muon events \label{fig:mu-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/muon_ed.png}}
  \hfill
  \subfloat[Proton events \label{fig:P-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/proton_ed.png}}
  \hfill
  \subfloat[Neutron events \label{fig:N-geo}]
  {\includegraphics[width=0.44\textwidth]{pics/neutron_ed.png}}
  \caption{The detector geometry with a 2 GeV particle events in three different materials. The top displays corresponds to lead, the middle displays to polypropylene, and the bottom displays to water. Red trajectories indicate negatively charged tracks, green neutral tracks, and blue positively charged tracks.}
  \label{fig:mean and std of nets}
\end{figure*}

\begin{appendix}

\section{Volumes in Detector Construction}\label{sec:AppendixA}

\end{appendix}

### 2015-04-04 p2_505/presentation/pres.tex
first (combining all the old repos and adding project 3)

\documentclass[mathserif,18pt,xcolor=table]{beamer}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bbm}
\usepackage{ulem}
\usepackage{slashed}
\usepackage{graphicx}
\usepackage{textpos}
\usepackage{listings}
\usepackage{epsfig}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage{enumerate}
\usepackage{fixltx2e}
\definecolor{dukeblue}{RGB}{0,0,156}
\definecolor{dukedarkblue}{RGB}{0,26,87}
\definecolor{dukeblack}{RGB}{79,79,79}
\definecolor{dukegray}{RGB}{79,79,79}
\definecolor{dukesecbrown}{RGB}{217,200,158}
\definecolor{dukesecblue}{RGB}{127,169,174}
\mode<presentation> {
  \usetheme{Boadilla}  
  \setbeamercovered{invisible}
  \setbeamertemplate{navigation symbols}{}  
  \setbeamertemplate{frametitle}[default][center]
  \setbeamerfont{title}{series=\bfseries,parent=structure}
  \setbeamerfont{frametitle}{series=\bfseries,parent=structure}
  \setbeamerfont{subtitle}{size=\scriptsize,series=\bfseries,parent=structure}
  \setbeamerfont{author}{size=\scriptsize,parent=structure}
  \setbeamerfont{institute}{size=\small,series=\bfseries,parent=structure}
  \setbeamerfont{date}{size=\scriptsize,parent=structure}
  \setbeamerfont{footline}{size=\tiny,parent=structure}
  \setbeamercolor{normal text}{bg=white,fg=dukeblack}
  \setbeamercolor{structure}{fg=dukeblue}
  \setbeamercolor{alerted text}{fg=red!85!black}
  \setbeamercolor{item projected}{use=item,fg=black,bg=item.fg!35}
  \setbeamercolor*{palette primary}{use=structure,fg=white, bg=dukeblue}
  \setbeamercolor*{palette secondary}{use=structure,bg=dukedarkblue,fg=white}
  \setbeamercolor*{framesubtitle}{fg=dukegray}
  \setbeamercolor*{block title}{parent=structure,fg=black,bg=dukeblue}
  \setbeamercolor*{block body}{fg=black,bg=dukeblack!10}
  \setbeamercolor*{block title alerted}{parent=alerted text,bg=black!15}
  \setbeamercolor*{block title example}{parent=example text,bg=black!15}
}

\makeatletter
\setbeamertemplate{footline}{
  \leavevmode
  \hbox{%
    \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{author in head/foot}%
      \usebeamerfont{author in head/foot}\insertshortauthor%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,center]{title in head/foot}%
      \usebeamerfont{title in head/foot}\insertshorttitle%
    \end{beamercolorbox}%
    \begin{beamercolorbox}[wd=.333333\paperwidth,ht=2.25ex,dp=1ex,right]{date in head/foot}%
      \usebeamerfont{date in head/foot}\insertshortdate{}\hspace*{2em}%
      \insertframenumber{} / \inserttotalframenumber\hspace*{2ex}%
  \end{beamercolorbox}}%
  \vskip0pt%
}
\makeatother

\AtBeginSection{\frame{\sectionpage}}

\defbeamertemplate{section page}{mine}[1][]{%
  \begin{centering}
    {\usebeamerfont{section name}\usebeamercolor[fg]{section name}#1}
    \vskip1em\par
    \begin{beamercolorbox}[sep=12pt,center]{part title}
      \usebeamerfont{section title}\insertsection\par
    \end{beamercolorbox}
  \end{centering}
}

\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{amsmath}
\renewcommand*{\thefootnote}{\fnsymbol{footnote}}
\pgfdeclareimage[height=1.7cm]{atlaslogo}{logos/atlas_logo.pdf}
\pgfdeclareimage[height=1.7cm]{dukelogo}{logos/dukelogo.pdf}
\title[PHY 505 Project 2]{Mesons Beyond \su3flavor \qqbar}
\author[DD, ME, JR, PZ]{{\small Douglas Davis, Matthew Epland, Justin Raybern, Pingchuan Zhao}}
\institute{\it{Duke University} \\ \mbox{} \\ \mbox{} \\ \pgfuseimage{dukelogo}}
\date[19 March 2014]{19 March 2014}

\DeclareRobustCommand{\su3flavor}{$\text{SU(3)}_{\text{flavor}}$}
\DeclareRobustCommand{\qqbar}{$q\overline{q}$}

\begin{document}

\beamertemplateballitem
\frame{\titlepage}
\addtobeamertemplate{frametitle}{}{}

\begin{frame}
  \frametitle{Outline}
  \begin{itemize}
  \item Introduction to non \su3flavor mesons.
  \item Theoretical states:
    \begin{itemize}
    \item QCD Hybrids
    \item Glueballs
    \item Tetraquarks
    \end{itemize}
  \item Discussion of the $\pi_1(1400)$ and $\pi_1(1600)$ exotics.
  \item Discussion of the $Z(4430)$ tetraquark.
  \item Summary
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Introduction}
  \begin{itemize}
  \item Mesons are defined as particle resonances with baryon number 0.
    \begin{align}
      B_n = \frac{1}{3}\left(n_q-n_{\overline{q}}\right) = 0
    \end{align}
  \item Therefore, mesons can be defined as any bound state containing and equal number of quarks and antiquarks (or none of each).
  \item We will discuss three types of mesons beyond \su3flavor \qqbar: QCD Hybrids, Glueballs, and Tetraquarks.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{QCD Hybrids}
  \begin{itemize}
  \item `Normal'  mesons are color--anticolor \su3flavor \qqbar~states.
    \begin{itemize}
    \item Gluon field binding the quarks implicitly assumed to be colorless, with vacuum quantum numbers.
    \end{itemize}
  \item Calculating \qqbar~meson quantum numbers:
    \begin{itemize}
    \item $S = 0, 1$; $L = 0, 1, 2, \ldots$; $J = L+S, L+S-1, \ldots, \left| L-S \right|$
    \item $P = (-1)^{L+1},~C = (-1)^{L+S}$
    \end{itemize}
  \item Allowed $J^{PC} = 0^{-+}, 0^{++}, 1^{--}, 1^{+-}, 1^{++}, 2^{--}, 2^{-+}, 2^{++},\ldots$ for (neutral) \su3flavor \qqbar~mesons \cite{ketzer}.
  \item Other $J^{PC}$'s such as $J^{PC} = 0^{--}, 0^{+-}, 1^{-+}, 2^{+-},\ldots$ are considered `exotic'.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{QCD Hybrids -- Continued}
  \begin{itemize}
  \item Full QCD approach allows for a wider variety of mesons:
    \begin{itemize}
    \item Gluon field can have a net color, opposite to the quark's net color.
    \item Gluon field can also take different quantum numbers.
    \end{itemize}
  \item Such mesons are called QCD hybrid's.
    \begin{itemize}
    \item Includes mesons with more than one \qqbar~pair.
    \end{itemize}
  \item Hybrids with allowed \su3flavor quantum numbers are difficult to detect experimentally; appear as an excess.
  \item Hybrids with exotic $J^{PC}$ quantum numbers easy to detect experimentally; ``smoking gun''.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{QCD Hybrids -- Continued}
  \begin{itemize}
  \item Non-perturbative/lattice QCD needed to calculate QCD hybrid's properties.
  \item Lattice QCD can work from first principles, but the results have large uncertainties and can depend on the specific methods used.
  \item Experimental measurements of QCD hybrids would help clarify the theoretical picture.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Glueballs}
  \begin{itemize}
  \item Gluons carry color charge and interact through the strong force.
  \item A glueball can be two or three (or more) gluons bound to one another in the absence of quarks.  It is color neutral.
    \begin{align}
      |gg\rangle, |ggg\rangle, |ggg\ldots\rangle
    \end{align}
  \item There are no experimentally confirmed glueballs – though there are several candidates.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Glueballs -- Continued}
  \begin{itemize}
  \item They exhibit significant mixing with mesons
  \item The masses of the lighest glueballs are predicted to be around the same mass as known standard mesons
  \item QCD sum rule is a popular technique for calculating glueball masses
  \end{itemize}
  \begin{center}
    \includegraphics[width=.725\linewidth]{../figures/glueballmass.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Tetraquarks}
  \begin{itemize}
  \item A 4 quark meson of the $|qq\overline{q}\overline{q}\rangle$ variety.
  \item Not two standard mesons interacting with one another (though this is how it might decay).
  \item Experimentally confirmed with $Z(4430)$.
  \item Higher numbers of quark groups have been theorized (pentaquarks, and more).
  \item Allowed by Lattice QCD.
  \end{itemize}
  \begin{center}
    \includegraphics[width=.7\linewidth]{../figures/tetra.pdf}
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{$\pi_1(1400)$ and $\pi_1(1600)$ -- Theoretical Prediction \& The Peripheral Experiments}
  \begin{itemize}
  \item Both the QCD String (QCD flux tube) Model and the Latice QCD model predict the hybrid exotic meson with $J^{PC}=1^{-+}$ around $1.9\pm 0.2$ GeV.
  \item Bag model predicts the exotic $J^{PC} = 1^{-+}$ at lower mass about $1.4$ GeV.
  \item Hybrid Mesons are expected to decay mainly into pairs of S- and P-wave mesons
  \item $\pi_1(1400)$, reported in the reaction $\pi^− p \rightarrow \eta \pi^− p$ at 18.3 GeV/$c$ by the E852 collaboration using the multi-particle Spectrometer (MPS) at the AGS.
  \item $\pi_1(1600)$, reported in the reaction $\pi^- p \rightarrow \pi^- \pi^- \pi^+ p$ at 18.3 GeV/$c$ by the E852
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{$\pi_1(1400)$ Peripheral Experimental Results}
  \begin{center}
    \includegraphics[width=.75\linewidth]{../figures/ping35.pdf}
  \end{center}
  \small{
    \begin{itemize}
    \item (L) Neutral parity exchange intensity. (R) Exotic parity exchange intensity.
    \item The peak in the $D_+$ amplitude is due to the $a_2(1320)$ meson, the peak in the $P_+$ amplitude is due to the exotic $\pi_1(1400)$. The exotic intensity is a small fraction (about 3\%) of the dominating $a_2(1320)$ contribution.
    \item Interation: $\pi^- p \rightarrow \eta\pi^-p$ (Intensity for the $\eta\pi^-$ system).
  \end{itemize}}
\end{frame}

\begin{frame}
  \frametitle{$\pi_1(1600)$ Peripheral Experimental Results}
  \begin{center}
    \includegraphics[width=.95\linewidth]{../figures/ping1600.pdf}
  \end{center}
  \small{
    \begin{itemize}
    \item $\pi_1(1600)$ decays into two pseudoscalar mesons. $\pi^- p\rightarrow \eta^{\prime} \pi^- p$. $\eta^{\prime}$ decays into $\eta$ and is reconstructed by $\eta\rightarrow\gamma\gamma$ The properties are consistent with the previous found exotic resonance $\pi_1(1600)$ from decay $\pi^- p \rightarrow \pi^- \pi^- \pi^+ p$.
  \end{itemize}}
\end{frame}

\begin{frame}
  \frametitle{$\pi_1(1400)$ and $\pi_1(1600)$ Peripheral and Annihilation Conclusions}
  \begin{center}
    \includegraphics[width=.9\linewidth]{../figures/pingtable.pdf}
  \end{center}
  \begin{itemize}
  \item There are also $p\overline{p}$ annihilation experiments evidence for $\pi_1(1400)$ as isovectors, not glueballs. Decays also in favor of hybrids.
    Mass is below theory predictions. Future examinations are ongoing.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{The $Z(4430)$ at $B$ Factories}
  \small{
    \begin{itemize}
    \item Evidence of a charged tetraquark candidate at approximately 4430 MeV was observed by the Belle Collaboration in 2007. BaBar could not confirm.
    \item Charged state ruled out simply excited charmonium state.
    \item Resonant structures from $B^0 \rightarrow \psi^{\prime}\pi K$ decay.
    \item $J^P$ was unknown at the time and other models were possible (not a definite non-$q\overline{q}$ meson). The mass is very close to mass of $D^+_1\overline{D}^{*0} + D^{*+}\overline{D}_1^0$. Suspected to possibly be a $D$ ``molecule.''
    \end{itemize}
    \begin{center}
      \includegraphics[width=.8\linewidth]{../figures/Z4430_diagram.pdf}
    \end{center}
  }
\end{frame}

\begin{frame}
  \frametitle{The $Z(4430)$ at the LHC}
  \small{
    \begin{itemize}
    \item In 2014 LHCb experiment confirmed the resonance at 13.9$\sigma$ and also confirmed $J^P$ of $1^+$, supporting the tetraquark candidate over the $D$ ``molecule'' candidate.
    \item Minimum quark content of $c\overline{c}d\overline{u}$.
    \item $m_{Z^{-}} = 4475 \pm 7^{+15}_{-25}$ MeV. $\Gamma_{Z^{-}} = 172 \pm 13^{+37}_{-34}$ MeV.
    \end{itemize}
    \begin{center}
      \begin{columns}
        \begin{column}{.5\linewidth}
          \begin{center}
            \includegraphics[width=1\linewidth]{../figures/Z4430_res.pdf}
          \end{center}
        \end{column}
        \begin{column}{.5\linewidth}
          \begin{center}
            \includegraphics[width=1\linewidth]{../figures/Z4430_jp.pdf}
          \end{center}          
        \end{column}
      \end{columns}
    \end{center}
  }
  \tiny{Figures from: \url{http://cds.cern.ch/record/1693968}}
\end{frame}

\begin{frame}
  \frametitle{Summary}
  \begin{itemize}
  \item We've discussed some mesons beyond \su3flavor \qqbar.
  \item Hybrids, Glueballs, and Tetraquarks are all Lattice QCD allowed states beyond the non-relativistic quark model
  \item Experimental evidence exists for non $q\overline{q}$ mesons, supporting non-perturbative QCD.
  \item This includes the exotic pions we discussed and the $Z(4430)$, a confirmed tetraquark.
  \end{itemize}
\end{frame}

\begin{thebibliography}{9}
  %\bibitem{Barnes}
  %  T. Barnes, \textit{Exotic Mesons, Theory and Experiment}, arXiv:hep-ph/0007296, \url{http://arxiv.org/abs/hep-ph/0007296}
\bibitem{ketzer}
  Bernhard Ketzer, \textit{Hybrid Mesons}, Proceedings of Science, Sixth International Conference on Quarks and Nuclear Physics, PoS(QNP2012)025, \url{http://pos.sissa.it/archive/conferences/157/025/QNP2012_025.pdf} %\url{http://arxiv.org/abs/1208.5125}
  %\bibitem{Olsen}
  %  Stephen Lars Olsen, \textit{QCD exotics}, Hyperfine Interactions, \textbf{229}, \url{http://dx.doi.org/10.1007/s10751-014-1061-4}
\bibitem{belle}
  S.~K.~Choi~\emph{et. al.}, \emph{Observation of a Resonancelike Structure in the} $\pi^{+-}\psi^{\prime}$ \emph{Mass Distribution in Exclusive} $B\rightarrow K \pi^{+-}\psi^{\prime}$ \emph{Decays}. Phys. Rev. Lett. \textbf{100}, 142001 8 April 2008.
\bibitem{z4430_dmolecule}
  E.~Braaten~and~M.~Lu, \emph{Line shapes of the Z(4430)} Phys. Rev. D \textbf{79} 051503(R) 11 March 2009
\bibitem{lhcb} % {j4}
  LHCb Collaboration, R. Aaij \textit{et al.}, \textit{Observation of the Resonant Character of the }$Z(4430{)}^{-}$\textit{ State}, Phys. Rev. Lett., \textbf{112}, 22 (2014). \url{http://dx.doi.org/10.1103/PhysRevLett.112.222002}%\url{http://link.aps.org/doi/10.1103/PhysRevLett.112.222002}

\bibitem{j1}
  K.A. Olive \textit{et al.} (Particle Data Group), \textit{Non--\qqbar~Mesons} (Revised March 2006 by C. Amsler), Chin. Phys. C, 38, 090001 (2014), \url{http://pdg.lbl.gov/2006/reviews/nonqqbar_mxxx050.pdf}
\bibitem{j2}
  Vincent Mathieu, Nikolai Kochelev, and Vicente Vento, \emph{The Physics of Glueballs}, Int. J. Mod. Phys. E \textbf{18}, 1 (2009), \url{http://www.worldscientific.com/doi/abs/10.1142/S0218301309012124}
\bibitem{j3}
  Wolfgang Ochs, \emph{The status of glueballs}, J. Phys. G: Nucl. Part. Phys. \textbf{40}, 043001 (2013), \url{http://dx.doi.org/10.1088/0954-3899/40/4/043001}

\end{thebibliography}
\end{document}

### 2015-04-04 p3_505/paper/paper2.tex
first (combining all the old repos and adding project 3)

--- 2015-04-04 p2_505/paper/paper2.tex
@@
 \usepackage{graphicx}
+\usepackage{fourier}
 \usepackage{subfig}
@@
 \usepackage{verbatim}
-
-%\usepackage[section]{placeins}
 \usepackage{float}
-
-\DeclareRobustCommand{\su3flavor}{$\text{SU(3)}_{\text{flavor}}$}
-\DeclareRobustCommand{\qqbar}{$q\overline{q}$}
 
@@
 
-\title{Mesons Beyond \su3flavor \qqbar}
+\title{Sterile Neutrinos}
 \author{Douglas Davis}
@@
 \begin{abstract}
-%\todo[inline]{Abstract goes here!}
-Mesons beyond \qqbar~are important for refining our understanding of QCD.
-Both hybrid and exotic mesons are discussed along with the theoretical motivation of, and experimental evidence for, each.
-The QCD hybrids would provide an excellent test of lattice QCD predictions.
-Glueballs and tetraquarks are exotic mesons allowed by QCD and there is experimental evidence of the latter and several experimental candidates for the former.
-There is much work to do but the results seem to point to an exciting future.
+an abstract
 \end{abstract}\maketitle
 
-\section{Theoretical Motivations}
-The particles we call mesons are usually introduced as consisting of a single \su3flavor quark-antiquark pair, \qqbar.
-This description is somewhat simplistic however as the formal definition of a meson is any hadron, a particle made up of quarks, that has baryon number $B=0$.
-\begin{align}
-\label{eq:1}
-B = \frac{1}{3}(n_{q}-n_{\overline{q}})
-\end{align}
+\section{theory}
+yep
 
-The $B=0$ definition allows us to group a larger number of theorized particles together under the name mesons; those that can not be fully described as \su3flavor \qqbar~states are typically called exotic mesons.
-In this paper we will be discussing three of the main types of exotic mesons: QCD Hybrids, Glueballs, and Tetraquarks.
-
-\subsection{QCD Hybrids}
-The usual \su3flavor \qqbar~mesons can be well described as being colorless bound states of $q_{i}\overline{q}_{j}$, $q_{i}$ with some color and $\overline{q}_{j}$ with the corresponding anticolor, and can be grouped into isospin triplets based on the constituent quark's flavors.
-Each meson can further be characterized by it's quantum numbers, namely it's isospin $I$, total angular momentum $J$, and parity $P$.
-Additionally, the neutrally charge state of each isospin triplet is an eigenstate of the charge conjugation operator $\hat{C}$ with eigenvalue $C$.
-$P$ and $C$ are functions (\ref{eq:2}) of the constituent quark's relative orbital angular momentum $L$ and their total intrinsic spin $S$ \cite{ketzer}.
-
-\begin{align}
-\label{eq:2}
-P = (-1)^{L+1},~C = (-1)^{L+S}
-\end{align}
-
-As quarks are spin--$\frac{1}{2}$ particles, the total spin can only take values of $S = 0\text{ or }1$, and as usual $L$ and $J$ can take values $L = 0, 1, 2, 3, \ldots$; $J = L+S, L+S-1, \ldots, \left| L-S \right|$.
-All together, this gives us allowed quantum numbers $J^{PC} = 0^{-+}, 0^{++}, 1^{--}, 1^{+-}, 1^{++}, 2^{--}, 2^{-+}, 2^{++},\ldots$ for the neutral \su3flavor \qqbar~mesons \cite{ketzer}.
-
-The quantum numbers that are not allowed for \su3flavor mesons, such as $J^{PC} = 0^{--}, 0^{+-}, 1^{-+}, 2^{+-},\ldots$ are commonly called exotic quantum numbers and are allowed when using Quantum Chromodynamics (QCD) to model the binding of the quarks in the meson.
-QCD can produce a wider variety of \qqbar~mesons, including those with exotic quantum numbers, because it takes into account the gluon field mediating the strong force between the two quarks.
-The gluon field can have quantum numbers of it's own that combine with the quark's in the mesons overall quantum numbers, thereby creating the possibility of mesons with exotic quantum numbers.
-
-In addition to the effect they have on a meson's potential quantum numbers, gluons carry color which allows for meson states to exist where the gluon field binding the quarks is not colorless as is implicitly assumed in the \su3flavor model.
-Such states are called hybrids.
-Mesons with multiple \qqbar~pairs are also considered QCD hybrids provided that the valence \qqbar's net color is still canceled by a colored gluon field \cite{ketzer}.
-
-As hybrid mesons are predicted by QCD, but not the \su3flavor model, they are an important area of study in order to experimentally verify the predictions of QCD.
-Hybrid mesons with non-exotic quantum numbers are experimentally hard to detect as they will only be apparent as an excess of states on top of what is predicted by \su3flavor.
-Exotic hybrid mesons are therefore an area of great experimental interest as it is relatively easy to measure a mesons $J^{PC}$ quantum numbers.
-
-In order to use quantum chromodynamics to predict the properties of hybrid mesons non-perturbative calculations must be performed.
-This is by and large done with the lattice method, i.e.
-lattice QCD.
-While lattice QCD in theory is able to predict the hybrid mesons properties from first principles, in practice the calculations produce results that have large uncertainties and depend on the exact methods used.
-Experimentally discovering exotic hybrid meson states, and precisely measuring their properties, is therefore an important test of QCD as well as the lattice methods used to derive it's results.
-
-\subsection{Glueballs}
-%\todo[inline]{Justin}
-A potential non--\qqbar~meson candidate is made from a collection of gluons bound to one another, or a glueball, in the absence of quarks.
-The key is that gluons carry color charge and feel the strong force just as quarks do \cite{j1}.
-Detecting glueballs, however, has proven to be incredibly challenging.
-They exhibit significant mixing with known mesons.
-Additionally, the masses of the lightest predicted glueballs, those with $J^{PC} = 0^{++} \text{and}~2^{++}$, are around the same mass as known standard mesons.
-There are several resonances that are potentially linked to glueballs but the evidence remains inconclusive.
-Many masses for glueballs can derived from theory using a variety of techniques with the QCD sum rule being one of the more popular approaches \cite{j2}.
-From QCD we would expect two gluon states to have $J \neq 1$, $C = +1$, $P = \pm1$.
-Three gluon states can have $C = \pm1$ \cite{j3}.
-
-\subsection{Tetraquarks}
-%\todo[inline]{Justin}
-Another exotic meson category is a tetraquark, or 4 quark meson.
-The $qq\overline{q}\overline{q}$~tetraquark is allowed by QCD, and unlike the glueball, has been experimentally detected with the $Z(4430)$ \cite{lhcb}.
-This should not be thought of as two standard mesons bound to one another (though that is a likely way the tetraquark would decay), but as two quarks and two antiquarks bound as one.
-In addition to tetraquarks, it has been theorized that there could exist pentaquarks ($qqqq\overline{q}$) and even higher numbers of quark groups.
-
-\section{Experimental Results to Date}
-
-\subsection{The $Z(4430)$ Tetraquark}
-In 2007, the Belle Collaboration (a $B$ factory experiment using the KEKB $e^+e^-$ linear collider at $\sqrt{s} = 10.58$ GeV, the $\Upsilon(4S)$ mass.) observed a distinct peak in the invariant mass distribution of $\pi^{\pm}\psi^{\prime}$ samples near 4430~MeV~\cite{belle}.
-The resonant peak was found in $B\rightarrow K \pi^{\pm}\psi^{\prime}$ decays.
-Figure~\ref{fig:z4430diag} shows the decay chain of such an interaction.
-
-\begin{figure}[H]
-  \centerline{\psfig{file=../figures/Z4430_diagram.pdf, width=0.95\linewidth}}
-  \caption{The creation and decay chain of the $Z(4430)$ tetraquark meson.}
-  \label{fig:z4430diag}
-\end{figure}
-
-Due to the charge of the observed resonance it was impossible for the candidate particle to simply be an exited charmonium state ($c\overline{c}$).
-Therefore, the observed resonance was a suspected tetraquark state (with quark content $c\overline{c}u\overline{d}$~or $c\overline{c}d\overline{u}$).
-The $J^P$ of the particle was not measured by Belle, and therefore other hypothesis different from the tetraquark could be made.
-Another possible state was the $D$ ``molecule.'' A superposition of $D$ meson states forming an isospin multiplet~\cite{z4430_dmolecule}:
-\begin{align}
-Z^+ &= \frac{1}{\sqrt{2}}\left(D^{*+}\overline{D}^0_1-D^+_1\overline{D}^{*0}\right), \\
-Z^0 &= \frac{1}{2}\left(D_1^0\overline{D}^{*0} - D^{*0}\overline{D}_1^0 + D_1^+D^{*-}-D^{*+}D^-_1\right),\\
-Z^- &= \frac{1}{\sqrt{2}}\left(D^0_1D^{*-} - D^{*0}D_1^-\right).
-\end{align}
-Another $B$ factory experiment, BABAR at SLAC, was unable to find the same resonant behavior. Therefore, the official tetraquark discovery was not confirmed by the $B$ factory experiments.
-
-In 2014, the LHCb collaboration at the LHC confirmed the existence of the tetraquark state $Z(4430)^-$ with quark content $c\overline{c}d\overline{u}$~\cite{lhcb}.
-LHCb confirmed the state at 13.9$\sigma$.
-The spin-parity of the state was measured to be $J^P = 1^+$, eliminating the possible $D$ molecule and confirming the tetraquark state.
-The mass and width measured by LHCb are:
-\begin{align}
-m_{Z^{-}} &= 4475 \pm 7^{+15}_{-25}\text{ MeV},\\
-\Gamma_{Z^{-}} &= 172 \pm 13^{+37}_{-34}\text{ MeV}.
-\end{align}
-Figures~\ref{fig:z4430lhcb1} and ~\ref{fig:z4430lhcb2} show the resonant peak and also the likelihood measurement of for the spin-parity.
-
-\begin{comment}
-\begin{figure*}[h!]
-  \centerline{\psfig{file=../figures/Z4430_res.pdf, width=.45\linewidth}
-  \psfig{file=../figures/Z4430_jp.pdf,width=.45\linewidth}}
-  \caption{LHCb plots}
-  \label{fig:z4430lhcb}
-\end{figure*}
-\end{comment}
-
-\begin{figure}[H]
-  \centerline{\psfig{file=../figures/Z4430_res.pdf, width=.95\linewidth}}
-  \caption{The LHCb measurement of the invariant mass of the $\pi\psi^{\prime}$ system. The peak corresponds to $m=4475$ MeV.}
-  \label{fig:z4430lhcb1}
-\end{figure}
-
- \begin{figure}[H]
-  \centerline{\psfig{file=../figures/Z4430_jp.pdf,width=.95\linewidth}}
-  \caption{The LHCb measurement of the spin-parity of the $Z(4430)$.}
-  \label{fig:z4430lhcb2}
- \end{figure}
-
-\subsection{$\pi(1400)$ and $\pi(1600)$}
-%\todo[inline]{Ping}
-
-Both the Lattice QCD model and the QCD string (QCD flux tube) models predict an exotic hybrid meson with $J^{PC} = 1^{-+}$ with a mass of around 1.9$\pm$0.2 GeV.
-However, the bag model predicts the exotic $1^{-+}$ at a lower mass of about 1.4 GeV.
-This exotic $1^{-+}$ meson decaying to $\eta \pi^{-}$, is called $\pi_{1}(1400)$.
-It is reported in the reaction $\pi^{-}p \rightarrow \eta \pi^{-}p$ at 18.3 GeV by the E852 collaboration using the Multi-Particle Spectrometer (MPS) at the AGS.
-$\eta$ is detected through $\gamma \gamma$ decay.
-Evidence of $\pi_{1}(1400)$ is also found in $\pi^{-}p \rightarrow \eta \pi^{0} n$ at 18.3 GeV at E852.
-
-\subsubsection{$p\overline{p}$~Annihilation}
-
-The crystal barrel collaboration at LEAR also searched for a $1^{−+}$ resonance in the $\eta \pi$ P-wave\footnote{Hybrid mesons are expected to predominantly decay into pairs of S and P-wave mesons.
-The S-wave, angular momentum quantum number $l=0$, corresponds to a classical head on collision.
-P, $l=1$ and D-waves, $l=2$ have a higher angular momentum between the initial and final states.} in low energy $p\overline{p}$~annihilation into $\pi\pi\eta$.
-The $\pi_{1}(1400)$ signal is weak in peripheral reactions and contains lots of backgrounds which mimic other 1400 resonances such as the $a_{2}(1320)$ meson; the signal is stronger in $p\overline{p}$~annihilation.
-Also, a similar $\pi_{1}(1600)$ is also reported both from many peripheral experiments and $p\overline{p}$~annihilation.
-A confirmation that they are the particle predicted by the flux tube and lattice QCD model requires further theoretical and experimental work.
-
-%\subsection{Implications for Lattice QCD}
-%\todo[inline]{??, we need some references on this!!!}
-
-\section{Conclusions}
-%\todo[inline]{Justin, Includes short discussion of exotic mesons Implications for Lattice QCD, could just be a rehash of what is said in the hybrids section}
-The study of exotic mesons will contribute a great deal to our understanding of QCD.
-In particular, work on hybrid mesons can eliminate poor lattice QCD models in favor of those that make better experimental predictions.
-Glueballs and tetraquarks remain elusive, though success in verifying the Z(4430) gives hope that others may soon follow.
-The field could be a fruitful area of research for some time.
-
-\newpage
+\subsection{a subsection question mark?}
+wut
 
 \begin{thebibliography}{9}
-%\bibitem{Barnes}
-%  T. Barnes, \textit{Exotic Mesons, Theory and Experiment}, arXiv:hep-ph/0007296, \url{http://arxiv.org/abs/hep-ph/0007296}
 \bibitem{ketzer}
   Bernhard Ketzer, \textit{Hybrid Mesons}, Proceedings of Science, Sixth International Conference on Quarks and Nuclear Physics, PoS(QNP2012)025, \url{http://pos.sissa.it/archive/conferences/157/025/QNP2012_025.pdf} %\url{http://arxiv.org/abs/1208.5125}
-%\bibitem{Olsen}
-%  Stephen Lars Olsen, \textit{QCD exotics}, Hyperfine Interactions, \textbf{229}, \url{http://dx.doi.org/10.1007/s10751-014-1061-4}
 \bibitem{belle}
@@
 \url{http://dx.doi.org/10.1103/PhysRevLett.112.222002}%\url{http://link.aps.org/doi/10.1103/PhysRevLett.112.222002}
-
 \bibitem{j1}
- K.A.
-Olive \textit{et al.} (Particle Data Group), \textit{Non--\qqbar~Mesons} (Revised March 2006 by C.
-Amsler), Chin.
-Phys.
-C, 38, 090001 (2014), \url{http://pdg.lbl.gov/2006/reviews/nonqqbar_mxxx050.pdf}
+ K.A. Olive \textit{et al.} (Particle Data Group), \textit{hi} (Revised March 2006 by C. Amsler), Chin. Phys. C, 38, 090001 (2014), \url{http://pdg.lbl.gov/2006/reviews/nonqqbar_mxxx050.pdf}
 \bibitem{j2}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Simulation and Results}
-\subsection{Simulation Setup}
-
-Modifications to the example for our project include changes to the \texttt{PrimaryGeneratorAction}, \texttt{PhysicsList} \texttt{DetectorConstruction}, \texttt{RunAction}, and \texttt{EventAction}.
-For the generation of particles, we change the energy of the incoming particle, and also vary the particle type.
-We use electrons, muons, protons, and neutrons.
-For the detector construction, we create what is essentially a single block of material, with eight independent ``channels.'' When the particle propagates through the detector, it is always going through the same material, but independent layer by layer hit information is created by declaring eight different \texttt{LogicalVolume} objects.\footnote{For further discussion of volumes in Geant4 see Appendix~\ref{sec:AppendixA}} At the beginning of the \texttt{RunAction}\footnote{For further discussion of run and event actions, see Appendix~\ref{sec:AppendixB}} a ROOT~\cite{ROOT} \texttt{TFile} is initialized along with a \texttt{TTree} for data output.
-Sixteen branches are created for storing the energy deposited and track length for each layer.
-In the \texttt{EventAction}, \texttt{G4HitsCollection} pointers are declared to retrieve the total energy deposited and track length in each layer.
-These objects are pulled from the \texttt{G4SensitiveDetector} Manager.
-A pointer to the \texttt{RunAction} is declared in the \texttt{EventAction} where the function \texttt{RunAction::UpdateTree} has the current event's energy deposition and track length values updated in the \texttt{RunAction}, filling the whole run's \texttt{TTree}.
-At the end of the number of requested events, the \texttt{RunAction::EndOfRunAction} function writes the ROOT \texttt{TTree} to the \texttt{TFile} and closes the \texttt{TFile}.
-
-\subsection{Event Visualizations and Qualitative Event Analysis}
-
-\begin{figure*}
-  \centering
-  \subfloat[Electron events \label{fig:e-geo}]
-  {\includegraphics[width=0.44\textwidth]{pics/electron_ed.png}}
-  \hfill
-  \subfloat[Muon events \label{fig:mu-geo}]
-  {\includegraphics[width=0.44\textwidth]{pics/muon_ed.png}}
-  \hfill
-  \subfloat[Proton events \label{fig:P-geo}]
-  {\includegraphics[width=0.44\textwidth]{pics/proton_ed.png}}
-  \hfill
-  \subfloat[Neutron events \label{fig:N-geo}]
-  {\includegraphics[width=0.44\textwidth]{pics/neutron_ed.png}}
-  \caption{The detector geometry with a 2 GeV particle events in three different materials.
-The top displays corresponds to lead, the middle displays to polypropylene, and the bottom displays to water.
-Red trajectories indicate negatively charged tracks, green neutral tracks, and blue positively charged tracks.}
-  \label{fig:mean and std of nets}
-\end{figure*}
-
-\begin{appendix}
-
-\section{Volumes in Detector Construction}\label{sec:AppendixA}
-
-\end{appendix}

### 2015-04-04 p1_505/example.cc

@@ -1,0 +1,161 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: example.cc 86065 2014-11-07 08:51:15Z gcosmo $
+//
+/// \file example.cc
+/// \brief Main program of the  example
+
+#include "DetectorConstruction.hh"
+#include "ActionInitialization.hh"
+
+#ifdef G4MULTITHREADED
+#include "G4MTRunManager.hh"
+#else
+#include "G4RunManager.hh"
+#endif
+
+#include "G4UImanager.hh"
+#include "G4UIcommand.hh"
+#include "FTFP_BERT.hh"
+#include "QGSP_BERT.hh"
+
+#include "Randomize.hh"
+
+#include "G4VisExecutive.hh"
+#include "G4UIExecutive.hh"
+
[…]

### 2015-04-04 p1_505/python/add_edep.py

@@ -1,0 +1,8 @@
+import sys
+import ROOT
+
+the_file = ROOT.TFile(sys.argv[1],'read')
+the_tree = the_file.Get('RunTree')
+
+for event in the_tree:
+    print event.L1_Edep + event.L2_Edep + event.L3_Edep + event.L4_Edep

### 2015-04-04 p1_505/python/edep_per_layer.py

@@ -1,0 +1,27 @@
+import ROOT
+import sys
+import looks
+
+looks.looks()
+
+the_file = ROOT.TFile(sys.argv[1],'read')
+the_tree = the_file.Get('RunTree')
+
+histo = ROOT.TH2D('histo','Electron in Water;E_{dep} (MeV);Layer;Events',30,1,0,8,0,8)
+
+for event in the_tree:
+    histo.Fill(event.L1_Edep,0)
+    histo.Fill(event.L2_Edep,1)
+    histo.Fill(event.L3_Edep,2)
+    histo.Fill(event.L4_Edep,3)
+    histo.Fill(event.L5_Edep,4)
+    histo.Fill(event.L6_Edep,5)
+    histo.Fill(event.L7_Edep,6)
+    histo.Fill(event.L8_Edep,7)
+
+c = ROOT.TCanvas()
+c.SetLogz()
+    
+histo.Draw('lego')
+
+raw_input('')

### 2015-04-04 p1_505/python/tracklength_electron.py

@@ -1,0 +1,50 @@
+import ROOT
+import looks
+looks.looks()
+
+file_F_Pb = ROOT.TFile('../data/FTFP_Pb_electron.root')
+file_F_PP = ROOT.TFile('../data/FTFP_PP_electron.root')
+file_F_H2 = ROOT.TFile('../data/FTFP_H2O_electron.root')
+
+tree_F_Pb = file_F_Pb.Get('RunTree')
+tree_F_PP = file_F_PP.Get('RunTree')
+tree_F_H2 = file_F_H2.Get('RunTree')
+
+h_F_Pb = ROOT.TH1D('h_F_Pb',';Avg. Track Length/Layer;N events',100,0,1000)
+h_F_PP = ROOT.TH1D('h_F_PP',';Avg. Track Length/Layer;N events',100,0,1000)
+h_F_H2 = ROOT.TH1D('h_F_H2',';Avg. Track Length/Layer;N events',100,0,1000)
+
+
+def filler(the_tree,the_hist):
+    for event in the_tree:
+        tlsum = event.L1_Tlen + event.L2_Tlen + event.L3_Tlen + event.L4_Tlen + event.L5_Tlen + event.L6_Tlen + event.L7_Tlen + event.L8_Tlen
+        tlavg = tlsum/8.0;
+        the_hist.Fill(tlavg)
+    the_hist.Scale(1./the_hist.Integral())
+        
+filler(tree_F_Pb,h_F_Pb)
+filler(tree_F_PP,h_F_PP)
+filler(tree_F_H2,h_F_H2)
+
+h_F_Pb.SetLineColor(ROOT.kBlack)
+h_F_PP.SetLineColor(ROOT.kRed)
+h_F_H2.SetLineColor(ROOT.kAzure-2)
+
+legend = ROOT.TLegend(.7,.7,.9,.9)
+legend.SetHeader('2 GeV e^{-}')
+legend.SetTextSize(0.042)
+legend.SetTextFont(42)
+legend.AddEntry(h_F_Pb,'Lead','l')
+legend.AddEntry(h_F_PP,'Poly','l')
+legend.AddEntry(h_F_H2,'Water','l')
+
+c = ROOT.TCanvas()
+c.SetLogy()
+
+h_F_PP.Draw()
+h_F_Pb.Draw('same')
+h_F_H2.Draw('same')
+
+legend.Draw('same')
+
+raw_input('')

### 2015-04-04 p1_505/python/tracklength_muon.py

@@ -1,0 +1,51 @@
+import ROOT
+import looks
+looks.looks()
+
+file_F_Pb = ROOT.TFile('../data/FTFP_Pb_muon.root')
+file_F_PP = ROOT.TFile('../data/FTFP_PP_muon.root')
+file_F_H2 = ROOT.TFile('../data/FTFP_H2O_muon.root')
+
+tree_F_Pb = file_F_Pb.Get('RunTree')
+tree_F_PP = file_F_PP.Get('RunTree')
+tree_F_H2 = file_F_H2.Get('RunTree')
+
+h_F_Pb = ROOT.TH1D('h_F_Pb',';Avg. Track Length/Layer;N events',100,40,200)
+h_F_PP = ROOT.TH1D('h_F_PP',';Avg. Track Length/Layer;N events',100,40,200)
+h_F_H2 = ROOT.TH1D('h_F_H2',';Avg. Track Length/Layer;N events',100,40,200)
+
+
+def filler(the_tree,the_hist):
+    for event in the_tree:
+        tlsum = event.L1_Tlen + event.L2_Tlen + event.L3_Tlen + event.L4_Tlen + event.L5_Tlen + event.L6_Tlen + event.L7_Tlen + event.L8_Tlen
+        tlavg = tlsum/8.0;
+        the_hist.Fill(tlavg)
+    the_hist.Scale(1./the_hist.Integral())
+        
+filler(tree_F_Pb,h_F_Pb)
+filler(tree_F_PP,h_F_PP)
+filler(tree_F_H2,h_F_H2)
+
+h_F_Pb.SetLineColor(ROOT.kBlack)
+h_F_PP.SetLineColor(ROOT.kRed)
+h_F_H2.SetLineColor(ROOT.kAzure-2)
+
+legend = ROOT.TLegend(.7,.7,.9,.9)
+legend.SetHeader('2 GeV #mu^{-}')
+legend.SetTextSize(0.042)
+legend.SetTextFont(42)
+legend.AddEntry(h_F_Pb,'Lead','l')
+legend.AddEntry(h_F_PP,'Poly','l')
+legend.AddEntry(h_F_H2,'Water','l')
+
+c = ROOT.TCanvas()
+
+h_F_PP.Draw()
+h_F_Pb.Draw('same')
+h_F_H2.Draw('same')
+
+legend.Draw('same')
+
+c.SaveAs('TL_muon.pdf')
+
+raw_input('')

### 2015-04-04 p1_505/python/tracklength_neutron.py

@@ -1,0 +1,51 @@
+import ROOT
+import looks
+looks.looks()
+
+file_F_Pb = ROOT.TFile('../data/FTFP_Pb_N.root')
+file_F_PP = ROOT.TFile('../data/FTFP_PP_N.root')
+file_F_H2 = ROOT.TFile('../data/FTFP_H2O_N.root')
+
+tree_F_Pb = file_F_Pb.Get('RunTree')
+tree_F_PP = file_F_PP.Get('RunTree')
+tree_F_H2 = file_F_H2.Get('RunTree')
+
+h_F_Pb = ROOT.TH1D('h_F_Pb',';Avg. Track Length/Layer;N events',50,0,100)
+h_F_PP = ROOT.TH1D('h_F_PP',';Avg. Track Length/Layer;N events',50,0,100)
+h_F_H2 = ROOT.TH1D('h_F_H2',';Avg. Track Length/Layer;N events',50,0,100)
+
+
+def filler(the_tree,the_hist):
+    for event in the_tree:
+        tlsum = event.L1_Tlen + event.L2_Tlen + event.L3_Tlen + event.L4_Tlen + event.L5_Tlen + event.L6_Tlen + event.L7_Tlen + event.L8_Tlen
+        tlavg = tlsum/8.0;
+        the_hist.Fill(tlavg)
+    the_hist.Scale(1./the_hist.Integral())
+        
+filler(tree_F_Pb,h_F_Pb)
+filler(tree_F_PP,h_F_PP)
+filler(tree_F_H2,h_F_H2)
+
+h_F_Pb.SetLineColor(ROOT.kBlack)
+h_F_PP.SetLineColor(ROOT.kRed)
+h_F_H2.SetLineColor(ROOT.kAzure-2)
+
+legend = ROOT.TLegend(.7,.7,.9,.9)
+legend.SetHeader('2 GeV n')
+legend.SetTextSize(0.042)
+legend.SetTextFont(42)
+legend.AddEntry(h_F_Pb,'Lead','l')
+legend.AddEntry(h_F_PP,'Poly','l')
+legend.AddEntry(h_F_H2,'Water','l')
+
+c = ROOT.TCanvas()
+c.SetLogy()
+
+h_F_PP.Draw()
+h_F_Pb.Draw('same')
+h_F_H2.Draw('same')
+
+legend.Draw('same')
+
+c.SaveAs('TL_neutron.pdf')
+raw_input('')

### 2015-04-04 p1_505/python/tracklength_proton.py

@@ -1,0 +1,50 @@
+import ROOT
+import looks
+looks.looks()
+
+file_F_Pb = ROOT.TFile('../data/FTFP_Pb_P.root')
+file_F_PP = ROOT.TFile('../data/FTFP_PP_P.root')
+file_F_H2 = ROOT.TFile('../data/FTFP_H2O_P.root')
+
+tree_F_Pb = file_F_Pb.Get('RunTree')
+tree_F_PP = file_F_PP.Get('RunTree')
+tree_F_H2 = file_F_H2.Get('RunTree')
+
+h_F_Pb = ROOT.TH1D('h_F_Pb',';Avg. Track Length/Layer;N events',100,0,200)
+h_F_PP = ROOT.TH1D('h_F_PP',';Avg. Track Length/Layer;N events',100,0,200)
+h_F_H2 = ROOT.TH1D('h_F_H2',';Avg. Track Length/Layer;N events',100,0,200)
+
+
+def filler(the_tree,the_hist):
+    for event in the_tree:
+        tlsum = event.L1_Tlen + event.L2_Tlen + event.L3_Tlen + event.L4_Tlen + event.L5_Tlen + event.L6_Tlen + event.L7_Tlen + event.L8_Tlen
+        tlavg = tlsum/8.0;
+        the_hist.Fill(tlavg)
+    the_hist.Scale(1./the_hist.Integral())
+        
+filler(tree_F_Pb,h_F_Pb)
+filler(tree_F_PP,h_F_PP)
+filler(tree_F_H2,h_F_H2)
+
+h_F_Pb.SetLineColor(ROOT.kBlack)
+h_F_PP.SetLineColor(ROOT.kRed)
+h_F_H2.SetLineColor(ROOT.kAzure-2)
+
+legend = ROOT.TLegend(.7,.7,.9,.9)
+legend.SetHeader('2 GeV p')
+legend.SetTextSize(0.042)
+legend.SetTextFont(42)
+legend.AddEntry(h_F_Pb,'Lead','l')
+legend.AddEntry(h_F_PP,'Poly','l')
+legend.AddEntry(h_F_H2,'Water','l')
+c = ROOT.TCanvas()
+
+h_F_PP.Draw()
+h_F_Pb.Draw('same')
+h_F_H2.Draw('same')
+
+legend.Draw('same')
+
+c.SaveAs('TL_proton.pdf')
+
+raw_input('')

### 2015-04-04 p1_505/src/CalorHit.cc

@@ -1,0 +1,92 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: CalorHit.cc 69586 2013-05-08 14:20:11Z gcosmo $
+//
+/// \file CalorHit.cc
+/// \brief Implementation of the CalorHit class
+
+#include "CalorHit.hh"
+#include "G4UnitsTable.hh"
+#include "G4VVisManager.hh"
+#include "G4Circle.hh"
+#include "G4Colour.hh"
+#include "G4VisAttributes.hh"
+
+#include <iomanip>
+
+G4ThreadLocal G4Allocator<CalorHit>* CalorHitAllocator = 0;
+
+//....oooOO0OOooo........oooOO0OOooo........oooOO0OOooo........oooOO0OOooo......
+
+CalorHit::CalorHit()
+  : G4VHit(),
+    fEdep(0.),
+    fTrackLength(0.)
+{}
+
[…]

### 2015-04-04 p1_505/src/CalorimeterSD.cc

@@ -1,0 +1,128 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: CalorimeterSD.cc 87359 2014-12-01 16:04:27Z gcosmo $
+//
+/// \file CalorimeterSD.cc
+/// \brief Implementation of the CalorimeterSD class
+
+#include "CalorimeterSD.hh"
+#include "G4HCofThisEvent.hh"
+#include "G4Step.hh"
+#include "G4ThreeVector.hh"
+#include "G4SDManager.hh"
+#include "G4ios.hh"
+
+//....oooOO0OOooo........oooOO0OOooo........oooOO0OOooo........oooOO0OOooo......
+
+CalorimeterSD::CalorimeterSD(
+			     const G4String& name, 
[…]

### 2015-04-04 p1_505/src/DetectorConstruction.cc

@@ -1,0 +1,445 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: DetectorConstruction.cc 87359 2014-12-01 16:04:27Z gcosmo $
+// 
+/// \file DetectorConstruction.cc
+/// \brief Implementation of the DetectorConstruction class
+
+#include "DetectorConstruction.hh"
+#include "CalorimeterSD.hh"
+#include "G4Material.hh"
+#include "G4NistManager.hh"
+
+#include "G4Box.hh"
+#include "G4LogicalVolume.hh"
+#include "G4PVPlacement.hh"
+#include "G4PVReplica.hh"
+#include "G4GlobalMagFieldMessenger.hh"
+#include "G4AutoDelete.hh"
+
+#include "G4SDManager.hh"
+
+#include "G4VisAttributes.hh"
+#include "G4Colour.hh"
+
[…]

### 2015-04-04 p1_505/src/EventAction.cc

@@ -1,0 +1,160 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: EventAction.cc 75604 2013-11-04 13:17:26Z gcosmo $
+// 
+/// \file EventAction.cc
+/// \brief Implementation of the EventAction class
+
+#include "EventAction.hh"
+#include "RunAction.hh"
+#include "CalorimeterSD.hh"
+#include "CalorHit.hh"
+
+#include "G4RunManager.hh"
[…]

### 2015-04-04 p1_505/src/PrimaryGeneratorAction.cc

@@ -1,0 +1,102 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: PrimaryGeneratorAction.cc 75215 2013-10-29 16:07:06Z gcosmo $
+// 
+/// \file PrimaryGeneratorAction.cc
+/// \brief Implementation of the PrimaryGeneratorAction class
+
+#include "PrimaryGeneratorAction.hh"
+
+#include "G4RunManager.hh"
+#include "G4LogicalVolumeStore.hh"
+#include "G4LogicalVolume.hh"
+#include "G4Box.hh"
+#include "G4Event.hh"
+#include "G4ParticleGun.hh"
+#include "G4ParticleTable.hh"
+#include "G4ParticleDefinition.hh"
+#include "G4SystemOfUnits.hh"
+#include "Randomize.hh"
+
[…]

### 2015-04-04 p1_505/src/RunAction.cc

@@ -1,0 +1,126 @@
+//
+// ********************************************************************
+// * License and Disclaimer                                           *
+// *                                                                  *
+// * The  Geant4 software  is  copyright of the Copyright Holders  of *
+// * the Geant4 Collaboration.  It is provided  under  the terms  and *
+// * conditions of the Geant4 Software License,  included in the file *
+// * LICENSE and available at  http://cern.ch/geant4/license .  These *
+// * include a list of copyright holders.                             *
+// *                                                                  *
+// * Neither the authors of this software system, nor their employing *
+// * institutes,nor the agencies providing financial support for this *
+// * work  make  any representation or  warranty, express or implied, *
+// * regarding  this  software system or assume any liability for its *
+// * use.  Please see the license in the file  LICENSE  and URL above *
+// * for the full disclaimer and the limitation of liability.         *
+// *                                                                  *
+// * This  code  implementation is the result of  the  scientific and *
+// * technical work of the GEANT4 collaboration.                      *
+// * By using,  copying,  modifying or  distributing the software (or *
+// * any work based  on the software)  you  agree  to acknowledge its *
+// * use  in  resulting  scientific  publications,  and indicate your *
+// * acceptance of all terms of the Geant4 Software license.          *
+// ********************************************************************
+//
+// $Id: RunAction.cc 87359 2014-12-01 16:04:27Z gcosmo $
+//
+/// \file RunAction.cc
+/// \brief Implementation of the RunAction class
+
+#include "RunAction.hh"
+//#include "Analysis.hh"
+
+#include "G4Run.hh"
+#include "G4RunManager.hh"
+#include "G4UnitsTable.hh"
[…]

### 2015-04-14 p3_505/paper/paper3.tex
skeletoning
paper updates doug

--- 2015-04-04 p3_505/paper/paper2.tex
@@
 \usepackage{graphicx}
-\usepackage{fourier}
 \usepackage{subfig}
@@
 \usepackage{float}
-
 \begin{document}
-
-\title{Sterile Neutrinos}
+\title{A Survey of Sterile Neutrinos}
 \author{Douglas Davis}
@@
 \begin{abstract}
-an abstract
+xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxx
 \end{abstract}\maketitle
+\section{Theory}
+\label{sec:theory}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\subsection{Theory Ping}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\subsection{Theory Justin}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\section{Previous Experimental Efforts}
+There exists considerable experimental evidence for the existence of a fourth neutrino at energy scales in the eV range.
+The sources for evidence include the short baseline neutrino oscillation experiments LSND (Liquid Scintillator Neutrino Detector) at Los Alamos~\cite{LSND}, and MiniBooNE (Mini Booster Neutrino Experiment) at Fermilab~\cite{mini1,mini2}.
+There is also the reactor neutrino anomaly~\cite{reactor_anom1}, and also in the radioactive source calibrations for the solar neutrino experiments GALLEX~\cite{gallex1,gallex2} and SAGE~\cite{sage1,sage2}.
+\subsection{LSND}
+The LSND experiment was developed to search for $\overline{\nu}_\mu \rightarrow \overline{\nu}_e$ oscillations.
+The source of $\overline{\nu}_\mu$ was the decay at rest (DAR) of $\mu^+$.
+LSND used a proton beam on a target to produce pions.
+The decay chain was as follows:
+\begin{align}
+  \begin{split}
+    \pi^+ &\rightarrow \mu^+  \nu_\mu, \\
+    \mu^+ &\rightarrow e^+  \nu_e  \overline{\nu}_\mu.
+  \end{split}
+\end{align}
+The $\overline{\nu}_e$ events were detected through the interaction:
+\begin{align}
+  \overline{\nu}_e  p \rightarrow e^+ n,
+\end{align}
+inside of a large Cherenkov detector filled with liquid scintillator, utilizing 1280 phototubes.
+Figure~\ref{fig:lsnd_miniboone}(L) shows the excess of $\overline{\nu}_e$ events observed by LSND.
+The observation was $87.9\pm 22.4\pm6$ events over expected background.
+Their final result claimed the existence of $\overline{\nu}_\mu\rightarrow\overline{\nu}_e$ oscillations at $\Delta m^2 \sim 1\text{ eV}$ with 3.8$\sigma$.
+This is in heavy conflict with the existing three light flavor neutrino model supported by atmospheric and solar neutrino oscillation experiments, as well as measurements at LEP predicting $N=2.984\pm0.0082$ for weakly interacting neutrino flavors.
+Therefore, at least one ``sterile'' neutrino is required.
+\begin{figure*}
+  \includegraphics[width=1\textwidth]{../figures/lsnd_miniboone.pdf}
+  \caption{(L) ... (C) ... (R) ...}
+  \label{fig:lsnd_miniboone}
+\end{figure*}
+\subsection{MiniBooNE Doug}
+The MiniBooNE experiment was developed to answer the LSND anomaly.
+\subsection{Reactor Anomaly}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\section{Current and Future Experimental Efforts}
+yep
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\subsection{Current Matt}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+experiment
+\subsection{Future Matt}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+is cool
 
-\section{theory}
-yep
-
-\subsection{a subsection question mark?}
-wut
+\section{Conclusions}
+xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+who cares
 
 \begin{thebibliography}{9}
-\bibitem{ketzer}
-  Bernhard Ketzer, \textit{Hybrid Mesons}, Proceedings of Science, Sixth International Conference on Quarks and Nuclear Physics, PoS(QNP2012)025, \url{http://pos.sissa.it/archive/conferences/157/025/QNP2012_025.pdf} %\url{http://arxiv.org/abs/1208.5125}
-\bibitem{belle}
-  S.~K.~Choi~\emph{et.
-al.}, \emph{Observation of a Resonancelike Structure in the} $\pi^{+-}\psi^{\prime}$ \emph{Mass Distribution in Exclusive} $B\rightarrow K \pi^{+-}\psi^{\prime}$ \emph{Decays}.
-Phys.
+\bibitem{LSND}
+  A.~Aguilar-Arevalo \emph{et al.} Evidence for neutrino oscillations from the observation of anti-neutrino(electron) appearance in a anti-neutrino(muon) beam.
+\emph{Phys.
 Rev.
-Lett.
-\textbf{100}, 142001 8 April 2008.
-\bibitem{z4430_dmolecule}
-  E.~Braaten~and~M.~Lu, \emph{Line shapes of the Z(4430)} Phys. Rev. D \textbf{79} 051503(R) 11 March 2009
-\bibitem{lhcb} % {j4}
-  LHCb Collaboration, R.
-Aaij \textit{et al.}, \textit{Observation of the Resonant Character of the }$Z(4430{)}^{-}$\textit{ State}, Phys.
+D.}, {\bf 64} 112007, 2001.
+\bibitem{mini1}
+  A.~Aguilar-Arevalo \emph{et al.} Event Excess in the MiniBooNE Search for $\nu_\mu \rightarrow \nu_e$ Oscillations. \emph{Phys. Rev. Lett.}, {\bf 105} 181801, 2010.
+\bibitem{mini2}
+  A.~Aguilar-Arevalo \emph{et al.} Improved Search $\nu_\mu \rightarrow \nu_e$ Oscillations in the MiniBooNE Experiment. \emph{Phys. Rev. Lett.}, {\bf 110} 161801, 2013.
+\bibitem{reactor_anom1}
+  G. Mention, M. Fechner, Th. Lasserre, Th.A. Mueller, D. Lhuillier, \emph{et al.} The Reactor Antineutrino Anomaly. \emph{Phys. Rev. D.}, {\bf 83} 073006, 2011.
+\bibitem{reactor_anom2}
+  Th. A. Mueller, D. Lhuillier, M. Fallot, A. Letourneau, S. Cormon, \emph{et al.} Improved Predictions of Reactor Antineutrino Spectra. \emph{Phys. Rev. C.}, {\bf 83} 054615, 2011.
+\bibitem{gallex1}
+  P.~Anselmann~\emph{et al.} First results from the Cr-51 neutrino source experiment with the GALLEX detector. \emph{Phys. Lett. B.}, {\bf 342} 440-450, 1995
+\bibitem{gallex2}
+  W.~Hampel~\emph{et al.} Final results of the Cr-51 neutrino source experiments in GALLEX. \emph{Phys. Lett. B.}, {\bf 420} 114-126, 1998.
+\bibitem{sage1}
+  Dzh.~N.~Abdurashitov, V.N.~Gavrin, S.V.
+Girin, V.V.
+Gorbachev, Tatiana V.
+Ibragimova, \emph{et al.} The Russian-American gallium experiment (SAGE) Cr neutrino source measurement.
+\emph{Phys.
 Rev.
-Lett., \textbf{112}, 22 (2014).
-\url{http://dx.doi.org/10.1103/PhysRevLett.112.222002}%\url{http://link.aps.org/doi/10.1103/PhysRevLett.112.222002}
-\bibitem{j1}
- K.A. Olive \textit{et al.} (Particle Data Group), \textit{hi} (Revised March 2006 by C. Amsler), Chin. Phys. C, 38, 090001 (2014), \url{http://pdg.lbl.gov/2006/reviews/nonqqbar_mxxx050.pdf}
-\bibitem{j2}
- Vincent Mathieu, Nikolai Kochelev, and Vicente Vento, \emph{The Physics of Glueballs}, Int.
-J.
-Mod.
-Phys.
-E \textbf{18}, 1 (2009), \url{http://www.worldscientific.com/doi/abs/10.1142/S0218301309012124}
-\bibitem{j3}
- Wolfgang Ochs, \emph{The status of glueballs}, J. Phys. G: Nucl. Part. Phys. \textbf{40}, 043001 (2013), \url{http://dx.doi.org/10.1088/0954-3899/40/4/043001}
+Lett.}, {\bf 77} 4708-4711, 1996.
+\bibitem{sage2}
+  J.~N.~Abdurashitov~\emph{et al.} Measurement of the response of the Russian-American gallium experiment to neutrinos from a Cr-51 source. \emph{Phys. Rev. C.}, {\bf 59} 2246-2263, 1999.
 
@@
 \end{document} %%% end of doc %%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-04-14 p3_505/paper/paper3.tex
added .gitignore and half of the MINOS section

@@
 \section{Current and Future Experimental Efforts}
-yep
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
-\subsection{Current Matt}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
-experiment
-\subsection{Future Matt}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
-is cool
+Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
+\subsection{Current}
+%Matt
+\subsubsection{MINOS/MINOS+}
+The MINOS (Main Injector Neutrino Oscillation Search) experiment ran from 2005 to 2012 and its upgraded second run, MINOS+, has taking data since September 2013.
+The MINOS experiment is a long baseline style detector with a neutrino beam source and near detector at Fermilab and a far detector located in northern Minnesota.
+Though it's primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data.
+MINOS detects sterile nutrinos via their disappearence in neutral-current (NC) and
+
+The latest results \cite{MINOS+} are still in preprint but
+
+\subsubsection{T2K}
+
+\subsubsection{Daya Bay}
+
+\subsection{Future}
+%Matt
+NOVA
+MicroBooNE
+DUNE (Formerly LBNE)
 
@@
 
+\bibitem{MINOS+}
+  Alexandre B. Sousa~\emph{(for the MINOS and MINOS+ Collaborations)} First MINOS+ Data and New Results from MINOS. arXiv:1502.07715v1 [hep-ex] \url{http://arxiv.org/abs/1502.07715}
+
 \end{thebibliography}

### 2015-04-14 p3_505/paper/paper3.tex
more for paper doug

@@
 \begin{figure*}
+  \centering
   \includegraphics[width=1\textwidth]{../figures/lsnd_miniboone.pdf}
-  \caption{(L) ... (C) ... (R) ...}
+  \caption{(L) From the LSND Experiment~\cite{LSND}.
+The expected number of events is represented by the red and green histograms, the blue histogram agrees with the data and models $\overline{\nu}_{\mu}\rightarrow \overline{\nu}_e$ oscillation.
+(C) The MiniBooNE electron event excess in neutrino and anti-neutrino sector~\cite{mini2}.
+(R) The MiniBooNE allowed regions in neutrino and anti neutrino mode, for a 2 neutrino model.
+Black stars represent the MiniBooNE best fit points for $\Delta m^2$ and $\sin^22\theta$.~\cite{mini2}}
   \label{fig:lsnd_miniboone}
 \end{figure*}
-\subsection{MiniBooNE Doug}
+\subsection{MiniBooNE}
 The MiniBooNE experiment was developed to answer the LSND anomaly.
+The MiniBooNE detector also used Cherenkov radiation with photomultiplier tubes as their tool for detecting neutrino interactions.
+The Booster neutrino beam at Fermilab was used to create a $\nu_\mu$ or $\overline{\nu}_\mu$ beam.
+Protons were accelerated to 8 GeV by the Booster accelerator and were then steered to a beryillium target where a magnetic focusing horn steered pions from the proton on target collisions down the beam line towards the detector; the pions would decay into muons and muon neutrinos between the horn and the detector.
+Changing the magnetic focusing horn mode allowed for a beam of mostly $\nu_\mu$ or mostly $\overline{\nu}_\mu$ (neutrino or anti neutrino modes).
+In neutrino mode $\pi^+$ were steered towards the detector, while $\pi^-$ were steered away; and the reverse for anti neutrino mode.
+This is to utilize the decays:
+\begin{align}
+  \begin{split}
+    \pi^+ &\rightarrow \mu^+\nu_\mu \\
+    \pi^- &\rightarrow \mu^-\overline{\nu}_\mu  
+  \end{split}
+\end{align}
+The full dataset from MiniBooNE was consistent with the LSND $\overline{\nu}_e$ excess, and therefore the sterile neutrino hypothesis survives with a best fit $\Delta m^2 \sim 1$ eV.
 \subsection{Reactor Anomaly}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+reactors are cool.
+
 \section{Current and Future Experimental Efforts}
@@
 \subsection{Current}
-%Matt
+% Matt
 \subsubsection{MINOS/MINOS+}
@@
 \subsection{Future}
-%Matt
+% Matt
 NOVA

### 2015-04-14 p3_505/paper/paper3.tex
small changes

--- 2015-04-14 p3_505/paper/paper3.tex
@@
 Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
-\subsection{Current}
+\subsection{Current Experiments}
 %Matt
@@
 Though it's primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data.
-MINOS detects sterile nutrinos via their disappearence in neutral-current (NC) and
+MINOS detects sterile neutrinos via their disappearance in neutral-current (NC) interactions, which are independent of the three known neutrino flavors and their oscillations, and charged-current (CC) interactions differing from those predicted from the mixing of the three neutrinos of the standard model.
 
-The latest results \cite{MINOS+} are still in preprint but
+The latest sterile neutrino results from MINOS \cite{MINOS} are still in preprint but
 
@@
 
-\subsection{Future}
+\subsection{Future Experiments}
 %Matt
@@
   J.~N.~Abdurashitov~\emph{et al.} Measurement of the response of the Russian-American gallium experiment to neutrinos from a Cr-51 source. \emph{Phys. Rev. C.}, {\bf 59} 2246-2263, 1999.
-
-\bibitem{MINOS+}
+\bibitem{MINOS}
   Alexandre B. Sousa~\emph{(for the MINOS and MINOS+ Collaborations)} First MINOS+ Data and New Results from MINOS. arXiv:1502.07715v1 [hep-ex] \url{http://arxiv.org/abs/1502.07715}

### 2015-04-14 p3_505/paper/paper3.tex

--- 2015-04-14 p3_505/paper/paper3.tex
@@
 Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
-\subsection{Current}
-% Matt
+\subsection{Current Experiments}
+%Matt
 \subsubsection{MINOS/MINOS+}
@@
 
-\subsection{Future}
-% Matt
+\subsection{Future Experiments}
+%Matt
 NOVA

### 2015-04-14 p3_505/paper/paper3.tex

--- 2015-04-14 p3_505/paper/paper3.tex
@@
 Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
-\subsection{Current}
-% Matt
+\subsection{Current Experiments}
+%Matt
 \subsubsection{MINOS/MINOS+}
-The MINOS (Main Injector Neutrino Oscillation Search) experiment ran from 2005 to 2012 and its upgraded second run, MINOS+, has taking data since September 2013.
+The MINOS (Main Injector Neutrino Oscillation Search) experiment ran from 2005 to 2012 and its wide-band fresh, MINOS+, has taking data since September 2013.
 The MINOS experiment is a long baseline style detector with a neutrino beam source and near detector at Fermilab and a far detector located in northern Minnesota.
 Though it's primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data.
-MINOS detects sterile nutrinos via their disappearence in neutral-current (NC) and
+MINOS detects sterile neutrinos via their disappearance in neutral-current (NC) interactions, which are independent of the three known neutrino flavors and their oscillations, and in charged-current (CC) interactions differing from those predicted from the mixing of the three neutrinos of the standard model.
+The latest sterile neutrino results from the collaboration \cite{MINOS} now exclude a large majority of the LSND allowed phase space, Figure~\ref{fig:MINOS}.
+These exclusion limits are predicted to increase by up to a factor of 2 as MINOS+ continues to run.
 
-The latest results \cite{MINOS+} are still in preprint but
+\begin{figure}[H]
+ \centering
+ \includegraphics[width=1\columnwidth]{../figures/minos2.pdf}
+ \caption{MINOS sterile neutrino exclusion limits~\cite{MINOS}. MINOS is now excluding (to the right of the red line) much of the LSND allowed phase space (shaded).}
+ \label{fig:MINOS}
+\end{figure}
 

### 2015-04-14 p3_505/presentation/pres.tex
first (combining all the old repos and adding project 3)
skeletoning
presentation blah
figures stuff and a little presentation
miniboone slide done

--- 2015-04-04 p2_505/presentation/pres.tex
@@
 \pgfdeclareimage[height=1.7cm]{dukelogo}{logos/dukelogo.pdf}
-\title[PHY 505 Project 2]{Mesons Beyond \su3flavor \qqbar}
+\title[PHY 505 Project 3]{Sterile Neutrinos, a Summary}
 \author[DD, ME, JR, PZ]{{\small Douglas Davis, Matthew Epland, Justin Raybern, Pingchuan Zhao}}
 \institute{\it{Duke University} \\ \mbox{} \\ \mbox{} \\ \pgfuseimage{dukelogo}}
-\date[19 March 2014]{19 March 2014}
-
-\DeclareRobustCommand{\su3flavor}{$\text{SU(3)}_{\text{flavor}}$}
-\DeclareRobustCommand{\qqbar}{$q\overline{q}$}
+\date[15 April 2014]{15 April 2014}
 
@@
   \begin{itemize}
-  \item Introduction to non \su3flavor mesons.
-  \item Theoretical states:
-    \begin{itemize}
-    \item QCD Hybrids
-    \item Glueballs
-    \item Tetraquarks
-    \end{itemize}
-  \item Discussion of the $\pi_1(1400)$ and $\pi_1(1600)$ exotics.
-  \item Discussion of the $Z(4430)$ tetraquark.
-  \item Summary
+  \item neutrinos are kewl
+  \item and maybe sterile who knows
   \end{itemize}
 \end{frame}
+
+\begin{frame}
+  \frametitle{The LSND Experiment}
+  \footnotesize{
+    \begin{columns}
+      \begin{column}{.5\linewidth}      
+        \begin{itemize}
+        \item Liquid Scintillator Neutrino Detector.
+        \item Los Alamos National Laboratory from 1993 to 1998.
+        \item Cherenkov detector using 1220 PMTs.
+        \item Protons strike target, use pions decaying at test to product $\overline{\nu}_\mu$.
+        \end{itemize}
+      \end{column}
+      \begin{column}{.5\linewidth}
+        \begin{itemize}
+        \item $117.9 \pm 22.4$ events, signal excess of $87.9\pm 22.4 \pm 6.0$.
+        \item $(\sin^2 2\theta,\Delta m^2)$ fit extracts allowed oscillation regions.
+        \item Best fits yields $\Delta m^2 \sim 1$ eV.
+        \item Disagrees with solar and atmospheric oscillation experiments and also LEP experiments for 3 light flavor neutrinos.
+        \end{itemize}
+      \end{column}
+    \end{columns}
+  }
+  \begin{center}
+    \includegraphics[width=.35\linewidth]{../figures/LSND_plot.pdf}
+  \end{center}
+\end{frame}
+\begin{frame}
+  \frametitle{The MiniBooNE Experiment}
+  \scriptsize{
+    \begin{columns}
+      \begin{column}{.5\linewidth}      
+        \begin{itemize}
+        \item Developed to answer the LSND anomaly.
+        \item Collected data at Fermilab from 2002 to 2012
+        \item Cherenkov detector using PMTs (just like LSND)
+        \item 8 GeV protons from the Booster strike beryllium target, pions decay producing neutrinos
+        \begin{center}
+          \includegraphics[width=.9\linewidth]{../figures/miniboone_excess.pdf}
+        \end{center}
+        \end{itemize}
+      \end{column}
+      \begin{column}{.5\linewidth}
+        \begin{itemize}
+        \item Consistent with the LSND anti neutrino oscillation result.
+          \begin{center}
+            \includegraphics[width=.6\linewidth]{../figures/miniboone_regions.pdf}
+          \end{center}
+        \item Detector systematics make it difficult to differentiate between $\pi^0\rightarrow\gamma\gamma$ and an electron in the detector.
+        \end{itemize}
+      \end{column}
+    \end{columns}
+  }
+\end{frame}
+
+\end{document}
+
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%% 
+%%%%%%%%%%%%%%%%%%
+
+% \DeclareRobustCommand{\su3flavor}{$\text{SU(3)}_{\text{flavor}}$}
+% \DeclareRobustCommand{\qqbar}{$q\overline{q}$}
 
@@
 \begin{thebibliography}{9}
-  %\bibitem{Barnes}
+  % \bibitem{Barnes}
   %   T. Barnes, \textit{Exotic Mesons, Theory and Experiment}, arXiv:hep-ph/0007296, \url{http://arxiv.org/abs/hep-ph/0007296}
@@
   Bernhard Ketzer, \textit{Hybrid Mesons}, Proceedings of Science, Sixth International Conference on Quarks and Nuclear Physics, PoS(QNP2012)025, \url{http://pos.sissa.it/archive/conferences/157/025/QNP2012_025.pdf} %\url{http://arxiv.org/abs/1208.5125}
-  %\bibitem{Olsen}
+  % \bibitem{Olsen}
   %   Stephen Lars Olsen, \textit{QCD exotics}, Hyperfine Interactions, \textbf{229}, \url{http://dx.doi.org/10.1007/s10751-014-1061-4}

### 2015-04-14 p3_505/paper/paper3.tex
'

@@
 The MINOS experiment is a long baseline style detector with a neutrino beam source and near detector at Fermilab and a far detector located in northern Minnesota.
-Though it's primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data.
+Though its primary physics goals lie in the area of neutrino oscillation measurements, MINOS and MINOS+ are now also running sterile neutrino analysis with their data.
 MINOS detects sterile neutrinos via their disappearance in neutral-current (NC) interactions, which are independent of the three known neutrino flavors and their oscillations, and in charged-current (CC) interactions differing from those predicted from the mixing of the three neutrinos of the standard model.

### 2015-04-14 p3_505/paper/paper3.tex
added citations

@@
 Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
-\subsection{Current Experiments}
+\subsection{Current Experimental Results}
 %Matt
@@
 
-\subsubsection{T2K}
+\subsubsection{Super-Kamiokande}
+
+~\cite{SuperK}
 
 \subsubsection{Daya Bay}
+
+~\cite{DayaBay}
 
@@
 \bibitem{MINOS}
-  Alexandre B. Sousa~\emph{(for the MINOS and MINOS+ Collaborations)} First MINOS+ Data and New Results from MINOS. arXiv:1502.07715v1 [hep-ex] \url{http://arxiv.org/abs/1502.07715}
+  Alexandre B. Sousa (for the MINOS and MINOS+ Collaborations) First MINOS+ Data and New Results from MINOS. arXiv:1502.07715v1 [hep-ex] \url{http://arxiv.org/abs/1502.07715}
+\bibitem{SuperK}
+  K.
+Abe~\emph{et al.} (Super-Kamiokande Collaboration) Limits on sterile neutrino mixing using atmospheric neutrinos in Super-Kamiokande.
+\emph{Phys.
+Rev.
+D}, {\bf 91}, 052019, 2015, \url{http://dx.doi.org/10.1103/PhysRevD.91.052019}
+\bibitem{DayaBay}
+ F. P.
+An~\emph{et al.} (Daya Bay Collaboration) Search for a Light Sterile Neutrino at Daya Bay.
+\emph{Phys.
+Rev.
+Lett.} {\bf 113}, 141802, 2014, \url{http://dx.doi.org/10.1103/PhysRevLett.113.141802}
 

### 2015-04-14 p3_505/paper/paper3.tex
doug updates to paper

--- 2015-04-14 p3_505/paper/paper3.tex
@@
 inside of a large Cherenkov detector filled with liquid scintillator, utilizing 1280 phototubes.
-Figure~\ref{fig:lsnd_miniboone}(L) shows the excess of $\overline{\nu}_e$ events observed by LSND.
+Figure~\ref{fig:lsnd_miniboone}L shows the excess of $\overline{\nu}_e$ events observed by LSND.
 The observation was $87.9\pm 22.4\pm6$ events over expected background.
@@
 (R) The MiniBooNE allowed regions in neutrino and anti neutrino mode, for a 2 neutrino model.
-Black stars represent the MiniBooNE best fit points for $\Delta m^2$ and $\sin^22\theta$.~\cite{mini2}}
+Black stars represent the MiniBooNE best fit points for $\Delta m^2$ and $\sin^22\theta$~\cite{mini2}.}
   \label{fig:lsnd_miniboone}
@@
 \end{align}
-The full dataset from MiniBooNE was consistent with the LSND $\overline{\nu}_e$ excess, and therefore the sterile neutrino hypothesis survives with a best fit $\Delta m^2 \sim 1$ eV.
+The full dataset from MiniBooNE was consistent with the LSND $\overline{\nu}_e$ excess, and therefore the sterile neutrino hypothesis survived.
+Figures~\ref{fig:lsnd_miniboone}C and \ref{fig:lsnd_miniboone}R show the electron excess and allowed oscillation regions from the MiniBooNE data, respectively.
+
+The systematics of the MiniBooNE detector made it difficult to differentiate between $\pi^0 \rightarrow \gamma\gamma$ and an electron in the detector; therefore, there is a large uncertainty in the total number of $\overline{\nu}_e$ events.
+It is also hypothesized that the neutrino flux from the Booster neutrino beam must be studied more rigorously.
 \subsection{Reactor Anomaly}
-reactors are cool.
+There has been an effort to re-analyse short baseline reactor neutrino oscillation data using newer flux calculations.
+These calculations show the posibility of the disappearane of reactor $\overline{\nu}_e$.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=1\columnwidth]{../figures/reactor_anom.png}
+  \caption{Plot for the short baseline reactor antineutrino anomaly.
+The red curve shows a possible three active neutrino mixing solution, using $\sin^2\theta_{13} = 0.06$.
+The blue curve shows a possible solution including a new neutrino mass state with $\Delta m^2 \gg 1\text{ eV}^2$.}
+\end{figure}
 
@@
 \subsection{Current Experiments}
-%Matt
+% Matt
 \subsubsection{MINOS/MINOS+}
@@
 \subsection{Future Experiments}
-%Matt
+% Matt
 NOVA

### 2015-04-14 p3_505/paper/paper3.tex

@@
 Due to the number of anomalous results to date sterile neutrino searches have become popular research topics at the majority of contemporary neutrino experiments, both those currently running and those in the design and construction phases.
+
 \subsection{Current Experiments}
-% Matt
+
+\subsection{Current Experimental Results}
+
 \subsubsection{MINOS/MINOS+}

### 2015-04-14 p3_505/paper/paper3.tex
matts changes
small changes

@@
 There has been an effort to re-analyse short baseline reactor neutrino oscillation data using newer flux calculations.
-These calculations show the posibility of the disappearane of reactor $\overline{\nu}_e$.
+These calculations show the possibility of the disappearance of reactor $\overline{\nu}_e$.
 \begin{figure}[H]
@@
 
+\begin{figure}[H]
+ \centering
+ \includegraphics[width=1\columnwidth]{../figures/sk1.pdf}
+ \caption{Super-Kamiokande sterile neutrino exclusion limits~\cite{SuperK}. TODODESCRIPTIONHERE}
+ \label{fig:SuperK}
+\end{figure}
+
 \subsubsection{Daya Bay}
@@
 ~\cite{DayaBay}
+
+\begin{figure}[H]
+ \centering
+ \includegraphics[width=1\columnwidth]{../figures/daya1.pdf}
+ \caption{Daya Bay sterile neutrino exclusion limits~\cite{DayaBay}. TODODESCRIPTIONHERE}
+ \label{fig:DayaBay}
+\end{figure}
 
@@
 xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
-who cares
 

### 2015-04-15 p3_505/presentation/pres.tex
added MINOS slide
matts changes
small changes
lots of updates to paper

@@
 \usepackage{ulem}
+\usepackage{feynmp-auto}
 \usepackage{slashed}
@@
   \begin{itemize}
-  \item neutrinos are kewl
-  \item and maybe sterile who knows
+  \item Introduction to neutrinos
+  \item Standard model neutrinos and neutrino oscillations
+  \item Sterile neutrinos
+  \item Past sterile evidence
+  \item Current and future status of sterile analysis.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Little Neutral One}
+  \begin{itemize}
+  \item No charge, small mass $\sim 0.320 \pm 0.081 \text{ eV}$ (sum of 3 flavors), spin 1/2 fermions.
+  \item Interactions mediated only by the weak interaction.
+  \item First theorized due to observed $\beta$-decay spectrum. Distribution of $\beta$ particle energy suggests three-body decay.
+  \item Existence of antineutrino is indicated from inverse $\beta$-decay.
+  \item Helicity: massless particle? \\
+    neutrinos: (always) left-handed from experiments.\\
+    antineutrinos: (always) right-handed from experiments.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Neutrinos in the Standard Model}
+  \begin{itemize}
+  \item In the standard model, exactly zero mass;
+    There is no mass term in the SM Lagrangian
+  \item Neutrinos and antineutrinos are distinct (Dirac particles)
+  \item Exactly three neutrinos with lepton number conservation
+  \item Couple to the $W$ and the $Z$ only.
+  \end{itemize}
+  \vspace{1cm}
+  \begin{columns}
+    \begin{column}{.5\textwidth}
+      \begin{center}
+        \begin{fmffile}{Znunu}
+          \begin{fmfgraph*}(60,50)
+            \fmfstraight
+            \fmfleftn{i}{1}\fmfrightn{o}{2}
+            \fmf{boson}{i1,v1}
+            \fmf{fermion}{o1,v1,o2}
+            \fmflabel{$Z$}{i1}
+            \fmflabel{$\overline{\nu}$}{o1}
+            \fmflabel{$\nu$}{o2}
+          \end{fmfgraph*}
+        \end{fmffile}
+      \end{center}
+    \end{column}
+    \begin{column}{.5\textwidth}
+      \begin{center}
+        \begin{fmffile}{Wenu}
+          \begin{fmfgraph*}(60,50)
+            \fmfstraight
+            \fmfleftn{i}{1}\fmfrightn{o}{2}
+            \fmf{boson}{i1,v1}
+            \fmf{fermion}{o1,v1,o2}
+            \fmflabel{$W^-$}{i1}
+            \fmflabel{$\overline{\nu}$}{o1}
+            \fmflabel{$e^-$}{o2}
+          \end{fmfgraph*}
+        \end{fmffile}
+      \end{center}
+    \end{column}
+  \end{columns}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Neutrinos beyond the Standard Model}
+  \begin{itemize}
+  \item Neutrino has mass as indicated by oscillations.
+  \item Helicity can therefore change, not permanent anymore.
+  \item There are many models and extensions added to the Standard Model to resolve questions.
+  \item Majorana particle and exact mass? $0\nu2\beta$ experiments are taking this on.
+  \item Where does the mass come from? See-Saw model is an idea.
+  \item CP violation and lepton number conservation? No if they are Majorana.
+  \item Sterile neutrinos, interacting only through gravitation?
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Flavor Oscillations}
+  \begin{itemize}
+  \item Mass eigenstates and flavor eigenstates are different; a unitary mixing matrix governs the transformation from one basis to the other.
+    \begin{align}
+      |\nu_j\rangle &= \sum_\alpha U_{\alpha j}|\nu_\alpha\rangle \\
+      |\nu_j(t)\rangle\rangle &= e^{-iE_j t}|\nu_j\rangle \\
+      |\nu_\alpha\rangle &= \sum_{j=1}U^{*}_{\alpha j}|\nu_j\rangle \\
+      \rightarrow |\nu_\alpha(t)\rangle &= \sum_{\beta=e,\mu,\tau}\sum_{j=1,2,3} U_{\beta j}U^*_{\alpha j}e^{-iE_jt}|\nu_\beta\rangle \\
+      P(\nu_\alpha\rightarrow \nu_\beta) &= |\langle\nu_\beta|\nu_\alpha\rangle|^2 \\
+      &= \sum_{j=1}^3\sum_{k=1}^3 U^*_{\alpha j}U_{\alpha k}U_{\beta j}U_{\beta k}^* e^{-i(E_j-E_k)k}.
+    \end{align}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Flavor Oscillations}
+  \begin{itemize}
+  \item Using $E\approx p$ (ultrarelativistic neutrinos), and $L = ct$:
+    \begin{align}
+      E_j - E_k &= \left(m_j^2 - m_k^2\right)/2E = \Delta m_{jk}^2/2E \\
+      P &= \sum_{j=1}^3\sum_{k=1}^3 U^*_{\alpha j}U_{\alpha k}U_{\beta j}U_{\beta k}^* e^{-i\Delta m_{jk}^2L/2E}
+    \end{align}
+  \item And therefore a complete expression without hte complex exponential is:
+    \begin{align}
+        P(\nu_{\alpha}\rightarrow\nu_{\beta}) = \delta_{\alpha\beta} & - 4\sum_{j>k}\Re\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin^2\left(\frac{\Delta m_{jk}^2}{4E}L\right) \nonumber \\
+  & + 2\sum_{j>k}\Im\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin\left(\frac{\Delta m_{jk}^2}{2E}L\right).
+    \end{align}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Flavor Oscillations}
+  \begin{itemize}
+  \item The Pontecorvo Maki, Nakagawa, Sakata Matrix
+\begin{align}
+  U & =
+  \begin{pmatrix}
+    U_{e1}    & U_{e2}    & U_{e3}   \\
+    U_{\mu1}  & U_{\mu2}  & U_{\mu3} \\
+    U_{\tau1} & U_{\tau3} & U_{\tau3}
+  \end{pmatrix} \\
+  \label{eq:umatrix}
+  & = \begin{pmatrix}
+    1 & 0 & 0 \\
+    0 & c_{23} & s_{23} \\
+    0 & -s_{23} & c_{23} \end{pmatrix} \times
+  \begin{pmatrix}
+    c_{13} & 0 & s_{13}e^{-i\delta} \\
+    0      & 1 & 0 \\
+    -s_{13}e^{i\delta} & 0 & c_{13} \end{pmatrix} \times
+  \begin{pmatrix}
+    c_{12} & s_{12} & 0 \\
+    -s_{12} & c_{12} & 0 \\
+    0 & 0 & 1
+  \end{pmatrix} \\
+  & = \begin{pmatrix}
+    c_{12}c_{13} & s_{12}c_{13} & s_{13}e^{-i\delta} \\
+    -s_{12}c_{23} - c_{12}s_{23}s_{13}e^{i\delta} & c_{12}c_{23} - s_{12}s_{23}s_{13}e^{i\delta} & s_{23}c_{13} \\
+    s_{12}s_{23} - c_{12}c_{23}s_{13}e^{i\delta} & -c_{12}s_{23} - s_{12}c_{23}s_{13}e^{i\delta} & c_{23}c_{13}
+  \end{pmatrix}.
+\end{align}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Flavor Oscillations}
+  \begin{itemize}
+    \item Simple two neutrino oscillation model.
+    \begin{align}
+      \begin{split}
+        \begin{pmatrix}
+          |\nu_\alpha\rangle \\
+          |\nu_\beta\rangle
+        \end{pmatrix} = 
+        \begin{pmatrix}
+          U_{\alpha 1} & U_{\alpha 2} \\
+          U_{\beta  1} & U_{\beta  2}
+        \end{pmatrix}
+        \begin{pmatrix}
+          |\nu_1\rangle \\
+          |\nu_2\rangle
+        \end{pmatrix}
+      \end{split}
+    \end{align}
+  \item Probability of oscillation from $\nu_\alpha$ to $\nu_\beta$:
+    \begin{align}
+      P(\nu_\alpha \rightarrow \nu_\beta) = \left|\langle \nu_\beta(x,t)|\nu_\alpha(0,0\rangle\right|^2
+    \end{align}
+  \item In this simple two neutrino model this yields probability expressions for the form:
+    \begin{align}
+      P(\nu_\mu\rightarrow\nu_e)   &= \sin^2(2\theta) \sin^2\left(\frac{1.27\Delta m^2[\text{eV}^2] L[\text{km}]}{E[\text{GeV}]}\right) \\
+      P(\nu_\mu\rightarrow\nu_\mu) &= 1 - \sin^2(2\theta) \sin^2\left(\frac{1.27\Delta m^2[\text{eV}^2] L[\text{km}]}{E[\text{GeV}]}\right)
+    \end{align}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Flavor Oscillations}
+  \begin{center}
+    \includegraphics[width=.6\linewidth]{../figures/oscillation_prob.pdf}    
+  \end{center}
+  \begin{itemize}
+\item Oscillation and survival probabilities for the two neutrino state with an initial $\nu_{\mu}$.
+The values for $\Delta m^2$ and $\sin^22\theta$ ($2.41\times 10^{-3}$ eV$^2$ and 0.97, respectively) are taken from 2013 MINOS result (PRL, 110, 251801)
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{See-Saw Model}
+  \begin{itemize}
+  \item Popular way to explain why there are light left-handed neutrinos and no observed right-handed neutrinos.
+  \item Right-handed states would be sterile -- don't interact (or interact only very weakly) via weak weak interactions.
+  \end{itemize}
+  \begin{center}
+    \includegraphics[width=.75\linewidth]{../figures/seesaw.pdf}
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{See-Saw continued}
+  \begin{align}
+    \begin{split}
+    \begin{pmatrix}
+      \overline{\nu} & \overline{N}
+    \end{pmatrix}
+    \begin{pmatrix}
+      0 & m_D \\
+      m_D & M
+    \end{pmatrix}
+    \begin{pmatrix}
+      \nu \\ N
+    \end{pmatrix}
+    \end{split}
+  \end{align}
+  \begin{itemize}
+    \item For the Dirac mass much ness than $M$, eigenvalues are $-m_D^2/M$ and $M$, with eigenstates made from mostly $\nu$ and $N$, respectively.
+    \item This means that as $M$ goes up or down $m_D^2/M$ does the opposite -- hence the See-Saw.
+    \item We are left with light-left handed neutrinos and heavy right-handed.
+    \item There are 3 types of See-Saw model (I, II, III) that involve the introduction of isospin singlets, scalar triplets, and fermion triplets.
+    \item Drawing distinctions is beyond the scope of this project.
   \end{itemize}
@@
   }
+\end{frame}
+
+\begin{frame}
+  \frametitle{The MINOS/MINOS+ Experiment}
+  \begin{itemize}
+  \item Long baseline, disappearance experiment
+  \item Latest results now exclude much of the LSND allowed phase space
+    \begin{center}
+      \includegraphics[width=.9\linewidth]{../figures/minos2.pdf}
+    \end{center}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{The Super-Kamiokande Experiment}
+  \begin{itemize}
+  \item Atmospheric, TODO experiment
+  \item Latest results TODO
+    \begin{center}
+      \includegraphics[width=.5\linewidth]{../figures/sk1.pdf}
+    \end{center}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{The Daya Bay Experiment}
+  \begin{itemize}
+  \item Reactor, TODO experiment
+  \item Latest results TODO
+    \begin{center}
+      \includegraphics[width=.5\linewidth]{../figures/daya1.pdf}
+    \end{center}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Conclusions}
+  \begin{itemize}
+  \item Sterile neutrinos\ldots
+  \end{itemize}
 \end{frame}

### 2015-04-15 p3_505/paper/paper3.tex
added two sections

@@
 \subsubsection{Super-Kamiokande}
-
-~\cite{SuperK}
+The Super-Kamiokande experiment is a 22.5 kton water Cherenkov detector that looks at atmospheric neutrino oscillations.
+Their latest sterile neutrino analysis~\cite{SuperK} used a sterile vacuum fit to put limits on the mixing between $\nu_{\mu}$ and a sterile neutrino $\nu_4$, parameterized by $\left|U_{\mu 4}\right|^2 \approx sin^2 \theta_{24}$, Figure~\ref{fig:SuperK}.
+Super-K can not measure $\Delta m^{2}_{41}$ in this analysis however, so when comparing to MiniBooNE the Super-K result is just a straight line.
+The Super-K result is is consistent with the MiniBooNE exclusion, providing verification on at least part of the MiniBooNE excluded region.
 
@@
  \includegraphics[width=1\columnwidth]{../figures/sk1.pdf}
- \caption{Super-Kamiokande sterile neutrino exclusion limits~\cite{SuperK}. TODODESCRIPTIONHERE}
+ \caption{Super-Kamiokande sterile neutrino exclusion limits at 90\% (solid) and 99\% (dashed) confidence~\cite{SuperK}.
+The Super-K sterile vacuum fit can not measure $\Delta m^{2}_{41}$ so it is plotted as a straight line when compared to the joint analysis of MiniBooNE and SciBooNE (dotted).}
  \label{fig:SuperK}
@@
 % Matt
-NOVA
-MicroBooNE
-DUNE (Formerly LBNE)
+While there are many future neutrino experiments in various stages of development, MicroBooNE and NOvA are the closest to producing new sterile neutrino results.
+As its name suggests, MicroBooNE is the larger follow up experiment to MiniBooNE.
+The MicroBooNE detector is an advanced time projection chamber with a 60 ton fiducial volume of liquid argon, and unlike MiniBooNE it will be able to distinguish between photons and electrons.
+It is widely expected that these significant improvements will allow MicroBooNE to resolve the MiniBooNE excess anomaly.
+The NOvA experiment is long baseline experiment that uses the same NuMI neutrino beam from Fermilab as MINOS and also has its far detector located in northern Minnesota.
+The 330 ton near and 14 kton far detectors are liquid scintillator designs located near the surface.
+NOvA's primary purpose is to make oscillation measurements, including potentially determining the ordering of the mass hierarchy, but will produce more sterile neutrino measurements as well.
 

### 2015-04-15 p3_505/paper/paper3.tex
Finished my sections, pasted in pings

@@
 xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
-\subsection{Theory Ping}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\subsection{Theory}
+%ping
+A neutrino is an electrically neutral, weakly interacting elementary particle with half integer spin.
+There are generations of neutrinos: electron-neutrino, muon-neutrino and tau-neutrino.
+It is first indicated from beta decay spectrum.
+From the experiment of radioactive decay, the outgoing electron energy varies differently, hence it is not a 2-body decay, but actually a 3-body decay.
+Pauli first called that small, neutral charge particle neutrino.
+Soon, this explanation is accepted among physicists and after a few years, anti-neutrinos is found from inverse beta decay.
+
+In the Standard Model, there are three generations of neutrinos.
+However, the Lagrangian has no neutrino mass term, giving neutrinos have exactly zero mass.
+Also, neutrinos and anti-neutrinos are distinct, separating by helicity.
+From experiment measurements, neutrinos always have left helicity and anti-neutrinos with right helicity.
+Neutrinos couple with W,Z bosons only in this model~\cite{ping1}.
+
+Although the Standard Model is exceedingly successful in explaining the interactions among fundamental particles, it can not be the ultimate theory, since it does not give explanation on gravity, dark matter, neutrino mass and neutrino oscillation.
+In 1960s, the solar neutrino experiment first indicates that there is a change of flavor when neutrinos traveling.
+This is also the first evidence that shows neutrinos have some mass.
+By using the detector Kamiokande and later the upgraded version, Super Kamiokande, they observe the flavor changes not only from solar neutrinos but also long-baseline reactor neutrinos.
+
+Starting in the late 1960s, there is a discrepancy using Standard Model to give explanations to neutrinos arriving from the sun~\cite{ping2}.
+The observed electron neutrinos are only between one third to one half of what Standard Model predicts.
+This problem remains unsolved until almost 30 years later as the observation of neutrino oscillation, which give a perfect explanation on solar neutrino problem.
+Since the neutrinos are changing flavors when traveling, the fraction of electron has changed to other flavors, and the fraction can be calculated from the mixing angles of oscillations.
+The mixing angles now have all been experimentally determined, with the last one $\theta_{13}$ found in 2012 in Daya Bay experiment~\cite{ping3}.
+However, the mixing angles and oscillation experiment only gives the relative mass between different flavors of neutrinos, this can not give the absolute mass of neutrinos.
+Hence the neutrino mass is crucial problem to understand nowadays.
+Neutrinos may be Majorana particles which is that they are their own antiparticles.
+If this is true, then not only the absolute mass of neutrinos can be observed, but also it explains why after the universe evolves, the matter dominates over anti-matter~\cite{ping4}.
+
+%The neutrinoless double beta decay experiment are very similar to dark matter search (WIMP search).
+The EXO experiment[6] is the Enriched Xenon Observatory in particle physics looking into detection of ``neutrino-less double beta decay''.
+It uses large amounts of xenon isotopically enriched in the isotope 136.
+A total 200-kg detector using liquid Xe is currently being installed at the Waste Isolation Pilot Plant (WIPP) near Carlsbad, New Mexico.
+“Many research and development efforts are underway for a ton-scale experiment, with the goal of probing new physics and the mass of the neutrino.” (EXO experiment).
+The CoGeNT[7] Dark Matter Experiment is a direct search of dark matter particles in a low-background germanium detector located at Soudan Underground Laboratory in Soudan, MN.
+The detector uses a single, 440-gram, high-purity germanium crystal cooled to liquid nitrogen temperatures in its measurements.
+The CoGeNT detector has very low energy threshold (~0.5 keV) so that it takes its advantage to search for nuclear recoil events from dark matter particles with relatively low mass.
+The detector is also built with very low background so that could recognize the signals very efficiently.
+Another advantage of the CoGent detector is that it senses only ionization charge from nuclear recoils and places limits on the mass and interaction cross-section of dark matter particles by excluding any candidate mass and cross-section pair that would result in a signal above the background of the detector.This experiment began collecting data in December 2009.
+
 \subsection{Theory Justin}
@@
 The Super-Kamiokande experiment is a 22.5 kton water Cherenkov detector that looks at atmospheric neutrino oscillations.
-Their latest sterile neutrino analysis~\cite{SuperK} used a sterile vacuum fit to put limits on the mixing between $\nu_{\mu}$ and a sterile neutrino $\nu_4$, parameterized by $\left|U_{\mu 4}\right|^2 \approx sin^2 \theta_{24}$, Figure~\ref{fig:SuperK}.
+Their latest sterile neutrino analysis~\cite{SuperK} used a sterile vacuum fit to put limits on the mixing between $\nu_{\mu}$ and a sterile neutrino $\nu_4$, parameterized by $\left|U_{\mu 4}\right|^2 \approx \sin^2 \theta_{24}$, Figure~\ref{fig:SuperK}.
 Super-K can not measure $\Delta m^{2}_{41}$ in this analysis however, so when comparing to MiniBooNE the Super-K result is just a straight line.
@@
 \subsubsection{Daya Bay}
-
-~\cite{DayaBay}
+The Daya Bay experiment is a reactor antineutrino experiment with six liquid scintillator detectors located at different baselines from six 2.9 GW$_\text{th}$ nuclear reactors.
+It is able to make sterile neutrino measurements at smaller mass squared differences, $10^{−3} \text{ eV}^2  \lesssim \left|\Delta m^{2}_{41}\right| \lesssim 0.1 \text{ eV}^2$, than have been probed by other experiments to date~\cite{DayaBay}.
+While they have many different combinations of their detectors data sets, the Daya Bay exclusion limits mark out more of the small mass difference phase space than the combined LSND and KARMEN limits.
 
@@
  \includegraphics[width=1\columnwidth]{../figures/daya1.pdf}
- \caption{Daya Bay sterile neutrino exclusion limits~\cite{DayaBay}. TODODESCRIPTIONHERE}
+ \caption{Daya Bay sterile neutrino exclusion limits~\cite{DayaBay}.}
  \label{fig:DayaBay}

### 2015-04-15 p3_505/paper/paper3.tex
fixed typos in my sections on the paper, added my slides.

@@
 
-\subsection{Current Experiments}
-
 \subsection{Current Experimental Results}
-
+%matt
 \subsubsection{MINOS/MINOS+}
@@
  \caption{Super-Kamiokande sterile neutrino exclusion limits at 90\% (solid) and 99\% (dashed) confidence~\cite{SuperK}.
-The Super-K sterile vacuum fit can not measure $\Delta m^{2}_{41}$ so it is plotted as a straight line when compared to the joint analysis of MiniBooNE and SciBooNE (dotted).}
+The Super-K sterile vacuum fit can not measure $\Delta m^{2}_{41}$ so it is plotted as a straight line when compared to the joint analysis of MiniBooNE and SciBooNE (dotted) at 90\% confidence.}
  \label{fig:SuperK}
@@
 It is widely expected that these significant improvements will allow MicroBooNE to resolve the MiniBooNE excess anomaly.
-The NOvA experiment is long baseline experiment that uses the same NuMI neutrino beam from Fermilab as MINOS and also has its far detector located in northern Minnesota.
+The NOvA experiment is a long baseline experiment that uses the same NuMI neutrino beam from Fermilab as MINOS and also has its far detector located in northern Minnesota.
 The 330 ton near and 14 kton far detectors are liquid scintillator designs located near the surface.

### 2015-04-15 p3_505/paper/paper3.tex
did abstract and some justin stuff

--- 2015-04-15 p3_505/paper/paper3.tex
@@
 \begin{abstract}
-  xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx xxxxxxxxxxxxxxxxxxxxx
+Neutrinos are light, weakly interacting, left-handed leptons in the Standard Model of particle physics.
+The discovery of neutrino oscillations (suggesting neutrinos actually have mass) has led to a fruitful time in neutrino research.
+Oscillation experiments measure the oscillation parameters $\Delta m^2$ and $\sin^2 2\theta$, which are the mass splittings of neutrino mass eigenstates and the mixing angles of a unitary transformaton matrix that governs the oscillation probabilities.
+Experiments have measured these parameters, and some results suggest the existence of the expected three weakly interacting neutrinos, but also the possibility of more neutrinos which are not weakly interacting, and are therefore given the name ``sterile'' neutrino.
+We summarize the theory and experimental evidence in the sterile neutrino sector.
 \end{abstract}\maketitle
@@
 \label{sec:theory}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
-\subsection{Theory Ping}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+\subsection{Standard Model and Beyond Standard Moden Neutrinos}
+A neutrino is an electrically neutral, weakly interacting elementary particle with half integer spin.
+There are three generations of leptons and their corresponding neutrino: electron \& electron-neutrino, muon \& muon-neutrino, and tau \& tau-neutrino.
+Neutrinos were first postulated as products in beta decay.
+From the experiment of radioactive decay, the outgoing electron energy varies; therefore it is not a 2-body decay, but actually a 3-body decay.
+Pauli first called that small, neutral charge particle neutrino.
+
+In the Standard Model, there are three generations of neutrinos, and the Lagrangian has no neutrino mass term.
+Also, neutrinos and anti-neutrinos are distinct, separated by helicity.
+From experiment measurements, neutrinos always have left helicity and anti-neutrinos with right helicity.
+Neutrinos couple ony with the $W$ and the $Z$ bosons in this model~\cite{ping1}.
+
+Although the Standard Model is exceedingly successful in explaining the interactions of fundamental particles, it can not be the ultimate theory, since it does not give explanations for gravity, dark matter, and now known neutrino mass, which is evident from neutrino oscillations.
+In the 1960s, the solar neutrino experiment first indicated that there is a change of flavor when neutrinos traveling.
+This is also the first evidence that shows neutrinos have mass (see section~\ref{sec:oscillate} for discussion of oscillations and mass).
+By using the detector Kamiokande and later the upgraded version, Super Kamiokande, they observe the flavor changes not only from solar neutrinos but also long-baseline reactor neutrinos and atmospheric neutrinos.
+
+Starting in the late 1960s, there is a discrepancy using Standard Model to give explanations to neutrinos arriving from the sun~\cite{ping2}.
+The observed electron neutrinos are only between one third to one half of what Standard Model predicts.
+This problem remains unsolved until almost 30 years later as the observation of neutrino oscillation, which give a perfect explanation on solar neutrino problem.
+Since the neutrinos are changing flavors when traveling, the fraction of electron has changed to other flavors, and the fraction can be calculated from the mixing angles of oscillations.
+The mixing angles now have all been experimentally determined, with the last one $\theta_{13}$ found in 2012 in Daya Bay experiment~\cite{ping3}.
+However, the mixing angles and oscillation experiment only gives the relative mass between different flavors of neutrinos, this can not give the absolute mass of neutrinos.
+Hence the neutrino mass is crucial problem to understand nowadays.
+Neutrinos may be Majorana particles which is that they are their own antiparticles.
+If this is true, then not only the absolute mass of neutrinos can be observed, but also it explains why after the universe evolves, the matter dominates over anti-matter~\cite{ping4}.
+
+\subsection{Oscillations}
+\label{sec:oscillate}
+Neutrino oscillations were theorized by Bruno Pontecorvo In the 1950's, when he believed neutrinos would oscillate between $\nu$ and $\overline{\nu}$ states similar to the mixing of neutral kaons $K^0$ and $\overline{K}^0$~\cite{ponte}.
+The theory that an electron neutrino could oscillate to a muon neutrino was first considered by Maki, Nakagawa, and Sakata~\cite{MNS}.
+
+In neutrino oscillation theory, neutrino mass eigenstates $|\nu_j\rangle$ are expressed as linear superposition of the flavor states, $|\nu_{\alpha}\rangle$, where $\alpha = e,\mu,\tau$:
+\begin{equation}
+  \left|\nu_j\right\rangle = \sum_{\alpha}U_{\alpha j}\left|\nu_{\alpha}\right\rangle,
+  \label{mass=flavor}
+\end{equation}
+where $U$ is a $3\times3$ unitary matrix called the PMNS (Pontecorvo, Maki, Nakagawa, Sakata) matrix.
+The matrix can be inverted to write the flavor eigenstates as a superposition of the mass eigenstates:
+\begin{equation}
+  \left|\nu_{\alpha}\right\rangle = \sum_{j=1}^3U^*_{\alpha j}\left|\nu_{j}\right\rangle.
+  \label{flavor=mass}
+\end{equation}
+Since the mass eigenstates have definite mass, $m_j$, and energy, $E_j$, they can be evolve in time as established by the Schr\"odinger Equation:
+\begin{equation}
+  \mathcal{H}\left|\nu_j(t)\right\rangle = i\hbar \frac{\partial}{\partial t}\left|\nu_j(t)\right\rangle,
+  \label{nu-schro1}
+\end{equation}
+yielding
+\begin{equation}
+  \left|\nu_j(t)\right\rangle = e^{-iE_jt}\left|\nu_j\right\rangle,
+  \label{nu-schro2}
+\end{equation}
+where $|\nu_j\rangle = |\nu_j(t=0)\rangle$. Combining with equation \ref{flavor=mass}:
+\begin{equation}
+  \left|\nu_{\alpha}(t)\right\rangle = \sum_{j=0}^3 U_{\alpha j}^* e^{-iE_jt}\left|\nu_j\right\rangle,
+  \label{flavor_whole}
+\end{equation}
+and now combining equation \ref{flavor_whole} with equation \ref{mass=flavor}:
+\begin{equation}
+  \left|\nu_{\alpha}(t)\right\rangle = \sum_{\beta=e,\mu,\tau}\sum_{j=0}^3U_{\beta j}U_{\alpha j}^* e^{-iE_jt}\left|\nu_{\beta}\right\rangle.
+\end{equation}
+The probability that a neutrino of flavor state $\nu_{\alpha}$ will oscillate into $\nu_{\beta}$ is written as:
+\begin{eqnarray}
+  P(\nu_{\alpha}\rightarrow\nu_{\beta}) &=& |\langle\nu_{\beta}|\nu_{\alpha}\rangle|^2 \nonumber \\
+                                        &=& \left|\sum_{j=0}^Ne^{-iE_jt}U_{\beta j}U_{\alpha j}^*\right|^2 \nonumber \\
+                                        &=& \sum_{j=1}^3\sum_{k=1}^3U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*e^{-i(E_j-E_k)t}.
+\end{eqnarray}
+In the ultra-relativistic limit (which neutrinos qualify for with $v \approx c$), we can write:
+\begin{equation}
+  E_i = \sqrt{m_i^2+p^2} \approx p + \frac{m_i^2}{2p} \\
+  \label{eq:approx1}
+\end{equation}
+\begin{equation}
+  E_j - E_k \approx p + \frac{m_j^2}{2p} - \left(p + \frac{m_k^2}{2p}\right),
+  \label{eq:approx2}
+\end{equation}
+and using $E \approx p$ (continuing with the ultra-relativistic limit):
+\begin{equation}
+  E_j - E_k = \frac{m_j^2 - m_k^2}{2E}.
+  \label{eq:approx3}
+\end{equation}
+Now we adopt the relation $\Delta m_{jk}^2 = m_j^2 - m_k^2$, and use $L = ct$, where $c\equiv1$,\footnote{we have been using $\hbar \equiv 1$ as well, adopting natural units for this derivation} to write:
+\begin{equation}
+  P(\nu_{\alpha}\rightarrow\nu_{\beta}) = \sum_{j=1}^3\sum_{k=1}^3U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*e^{-i\Delta m_{jk}^2L/2E}.
+\end{equation}
+We can now write a complete expression without a complex exponent as:
+\begin{widetext}
+  \begin{align}
+    \begin{split}
+      P(\nu_{\alpha}\rightarrow\nu_{\beta}) = \delta_{\alpha\beta} & - 4\sum_{j>k}\Re\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin^2\left(\frac{\Delta m_{jk}^2}{4E}L\right) \\
+      & + 2\sum_{j>k}\Im\left(U_{\alpha j}^*U_{\alpha k}U_{\beta j}U_{\beta k}^*\right)\sin\left(\frac{\Delta m_{jk}^2}{2E}L\right).
+    \end{split}
+  \end{align}
+\end{widetext}
+  We can write the unitary PMNS matrix as~\cite{matrixpara}, where $\delta$ is the neutrino sector $CP$ violating phase:
+\begin{widetext}
+  \begin{align}
+    \begin{split}
+      U & =
+      \begin{pmatrix}
+        U_{e1}    & U_{e2}    & U_{e3}   \\
+        U_{\mu1}  & U_{\mu2}  & U_{\mu3} \\
+        U_{\tau1} & U_{\tau3} & U_{\tau3}
+      \end{pmatrix} \\
+      \label{eq:umatrix}
+      & = \begin{pmatrix}
+        1 & 0 & 0 \\
+        0 & c_{23} & s_{23} \\
+        0 & -s_{23} & c_{23} \end{pmatrix} \times
+      \begin{pmatrix}
+        c_{13} & 0 & s_{13}e^{-i\delta} \\
+        0      & 1 & 0 \\
+        -s_{13}e^{i\delta} & 0 & c_{13} \end{pmatrix} \times
+      \begin{pmatrix}
+        c_{12} & s_{12} & 0 \\
+        -s_{12} & c_{12} & 0 \\
+        0 & 0 & 1
+      \end{pmatrix} \\
+      & = \begin{pmatrix}
+        c_{12}c_{13} & s_{12}c_{13} & s_{13}e^{-i\delta} \\
+        -s_{12}c_{23} - c_{12}s_{23}s_{13}e^{i\delta} & c_{12}c_{23} - s_{12}s_{23}s_{13}e^{i\delta} & s_{23}c_{13} \\
+        s_{12}s_{23} - c_{12}c_{23}s_{13}e^{i\delta} & -c_{12}s_{23} - s_{12}c_{23}s_{13}e^{i\delta} & c_{23}c_{13}
+      \end{pmatrix}.
+    \end{split}
+  \end{align}
+\end{widetext}
+In a simplified model, for example a two neutrino case, we can write the mixing matrix as:
+\begin{equation}
+  U = \begin{pmatrix}
+    \cos\theta & \sin\theta \\
+    -\sin\theta & \cos\theta
+  \end{pmatrix}.
+\end{equation}
+If we consider a two neutrino case consisting of $\nu_{\mu}$'s and $\nu_e$'s, we can write:
+\begin{equation}
+  \begin{pmatrix}
+    \nu_e \\
+    \nu_{\mu}
+  \end{pmatrix} = \begin{pmatrix}
+    \cos\theta & \sin\theta \\
+    -\sin\theta & \cos\theta
+  \end{pmatrix}
+  \begin{pmatrix}
+    \nu_1 \\
+    \nu_2
+  \end{pmatrix}
+\end{equation}
+\begin{eqnarray}
+  \left|\nu_{\mu}(t)\right\rangle &=& -\sin\theta e^{-iE_1t}\left|\nu_1\right\rangle+\cos\theta e^{-iE_2t}\left|\nu_2\right\rangle \\
+  \left|\nu_e\right\rangle &=& \cos\theta\left|\nu_1\right\rangle+\sin\theta\left|\nu_2\right\rangle
+\end{eqnarray}
+Utilizing equations \ref{eq:approx1}, \ref{eq:approx2}, \ref{eq:approx3} and adopting $L=t$ once again:
+\begin{equation}
+  P(\nu_{\mu}\rightarrow\nu_e) = |\langle \nu_e | \nu_{\mu}(t)\rangle|^2 = \sin^2 2\theta \sin^2\left(\frac{\Delta m^2 L}{4E}\right).
+\end{equation}
+Conversion from natural units to units common to the laboratory, we arrive at a probability equation of the form:
+\begin{equation}
+  P(\nu_{\mu}\rightarrow\nu_e) = \sin^22\theta\sin^2\left(\frac{1.27\Delta m^2 \left[\text{eV}^2\right] L [\text{km}]}{E[\text{GeV}]}\right).
+  \label{eq:osc}
+\end{equation}
+Naturally, the probability of survival is defined as:
+\begin{equation}
+  P(\nu_{\mu}\rightarrow\nu_{\mu}) = 1-\sin^22\theta\sin^2\left(\frac{1.27\Delta m^2 \left[\text{eV}^2\right] L [\text{km}]}{E[\text{GeV}]}\right).
+  \label{eq:surv}
+\end{equation}
 \subsection{Theory Justin}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+The See-Saw method is the most common way of explaining the possible existence of sterile neutrinos.
+The fact that there are no right-handed neutrinos, despite there being both left- and right-handed leptons, leaves theorists unsatisfied.
+Experiment confirms the lack of right-handed neutrinos.
+The small masses of the neutrinos is also unsettling.
+This can be reconciled by introducing right-handed terms in the neutrino Lagrangian~\cite{justin1}.
+The Lagrangian for a neutrino should include a mass term.
+The idea of sterile neutrinos arises from the addition of some nonzero right-handed component in the Lagrangian.
+This is seen in
+\begin{align}
+  \begin{split}
+  \begin{pmatrix}
+    \overline{\nu} & \overline{N}
+  \end{pmatrix}
+  \begin{pmatrix}
+    0 & m_D \\ m_D & M
+  \end{pmatrix}
+  \begin{pmatrix}
+    \nu \\ N
+  \end{pmatrix},
+  \end{split}
+\end{align}
+where $\nu$ and $N$ are the left- and right-handed neutrino components, respectively.
+The Dirac mass is given by $m_D$ and the Majorana mass is given my $M$.
+It is easy to show that the eigenvalues of this matrix, if $m_D \ll M$, are $m_D^2/M$ and $M$ with the approximate eigenvectors associated with the left- and right- handed neutrinos.
+This means that we have the normal neutrino eigenstate and a sterile neutrino eigenstate.
+
 \section{Previous Experimental Efforts}
@@
   \includegraphics[width=1\columnwidth]{../figures/daya1.pdf}
- \caption{Daya Bay sterile neutrino exclusion limits~\cite{DayaBay}. TODODESCRIPTIONHERE}
+  \caption{Daya Bay sterile neutrino exclusion limits~\cite{DayaBay}.}
   \label{fig:DayaBay}
@@
 \begin{thebibliography}{9}
+\bibitem{ping1}
+  T2K Experiment -- \url{http://t2k-experiment.org/neutrinos/in-the-standard-model/}
+
+\bibitem{ponte}
+  B.~Pontecorvo,
+  Mesonium and antimesonium,
+  \emph{Sov. Phys. JETP} {\bf 6}, 429 (1957).
+\bibitem{MNS}
+  Z.~Maki, M.~Nakagawa, and S.~Sakata,
+  Remarks on the Unified Model of Elementary Particles,
+  \emph{Progress of Theoretical Physics} {\bf 28}(5), 870–880 (1962).
+\bibitem{matrixpara}
+  L.~L.~Chau and W.~Y.~Keung,
+  Comments on the Parametrization of the Kobayashi-Maskawa Matrix,
+  \emph{Phys. Rev. Lett.}, {\bf 53}, 1802 (1984).
+
+\bibitem{justin1}
+Biswajoy Brahmachari, A see-saw model of sterile neutrino, \emph{Physics Letters B}, {\bf 461} 3, 26 August 1999, Pages 243-247.
 \bibitem{LSND}

### 2015-04-15 p3_505/presentation/pres.tex

@@
         \item Developed to answer the LSND anomaly.
-        \item Collected data at Fermilab from 2002 to 2012
-        \item Cherenkov detector using PMTs (just like LSND)
-        \item 8 GeV protons from the Booster strike beryllium target, pions decay producing neutrinos
+        \item Collected data at Fermilab from 2002 to 2012.
+        \item Cherenkov detector using PMTs (just like LSND).
+        \item 8 GeV protons from the Booster strike beryllium target, pions decay producing neutrinos.
           \begin{center}
@@
         \begin{itemize}
-        \item Consistent with the LSND anti neutrino oscillation result.
+        \item Consistent with the LSND antineutrino oscillation result.
           \begin{center}
@@
   \begin{itemize}
-  \item Atmospheric, TODO experiment
-  \item Latest results TODO
+  \item Atmospheric neutrino oscillation experiment, 22.5 kton water Cherenkov detector
+  \item Can not measure $\Delta m^{2}_{41}$ in this analysis
+  \item Super-K exclusion limits at 90\% (solid) and 99\% (dashed) confidence
+  \item Joint analysis of MiniBooNE and SciBooNE (dotted) at 90\% confidence
     \begin{center}
@@
   \begin{itemize}
-  \item Reactor, TODO experiment
-  \item Latest results TODO
+  \item Reactor antineutrino experiment
+  \item Six liquid scintillator/PMTs detectors, six nuclear reactors at different baselines
+  \item Sterile neutrino measurements possible at smaller mass squared differences than other experiments, $10^{−3} \text{ eV}^2  \lesssim \left|\Delta m^{2}_{41}\right| \lesssim 0.1 \text{ eV}^2$
     \begin{center}
@@
     \end{center}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}
+  \frametitle{(Selected) Future Experiments}
+  \begin{itemize}
+  \item MicroBooNE: Larger follow up experiment to MiniBooNE
+   \begin{itemize}
+     \item Time projection chamber with a 60 ton fiducial volume of liquid argon
+     \item Will be able to distinguish between photons and electrons
+     \item Expected to resolve the MiniBooNE excess anomaly
+   \end{itemize}
+  \item NOvA: Long baseline experiment using the NuMI beam
+   \begin{itemize}
+     \item 330 ton near and 14 kton far liquid scintillator detectors located near the surface
+     \item Primary physics goals are oscillation measurements, will also perform sterile neutrino measurements
+   \end{itemize}
   \end{itemize}

### 2015-04-15 p3_505/presentation/pres.tex
added the conclusion slide

@@
   \begin{itemize}
-  \item Sterile neutrinos\ldots
+  \item A discovery of heavy, right-handed sterile neutrinos would have a large impact on our understanding of neutrino physics
+  \item Some, very contested, evidence for sterile neutrinos has been found at LSND and MiniBooNE
+  \item MINOS+, Super-K, Daya Bay, MicroBooNE, and NoVA will continue searching for sterile neutrinos
+  \item Expect to see the anomalies be confirmed or refuted in the next year or two
   \end{itemize}

### 2015-04-15 p3_505/paper/paper3.tex
done?

@@
 \end{abstract}\maketitle
+\section{Introduction}
+The Standard Model of Particle Physics continues to succeed wonderfully in making predictions in nature.
+The discovery of the Higg's Boson is just one example of the predictive power of the model.
+There are places in nature, however, where the Standard Model fails - the fact that neutrinos have mass, which is confirmed by flavor oscillations, is completely outside the model.
+The smallness of the neutrino mass cannot be explained definitively by theory either.
+And even more interestingly, neutrinos seem to only come as left-handed particles.
+
+In this work we lay out some of basic Standard Model theory concerning neutrinos.
+We then describe oscillations and the significance of that behavior.
+We then explain sterile neutrinos - neutrinos that do not interact via the weak force, and how the see-saw model of neutrino mass can simultaneously explain light, left-handed neutrinos.
+We discuss some of the contested experimental evidence for sterile neutrinos and future experimental work in the field.
 \section{Theory}
 \label{sec:theory}
-\subsection{Standard Model and Beyond Standard Moden Neutrinos}
+\subsection{Standard Model, and Beyond, Neutrinos}
 A neutrino is an electrically neutral, weakly interacting elementary particle with half integer spin.
@@
 \end{equation}
-\subsection{Theory Justin}
+Oscillation and survival probabilities for $\nu_\mu\rightarrow\nu_e$ are shown in Figure~\ref{fig:probs}.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=1\linewidth]{../figures/oscillation_prob.pdf}
+  \caption{Oscillation and survival probabilities for the two neutrino state with an initial $\nu_{\mu}$.
+The values for $\Delta m^2$ and $\sin^22\theta$ ($2.41\times 10^{-3}$ eV$^2$ and 0.97, respectively) are taken from 2013 MINOS result (PRL, 110, 251801)}
+  \label{fig:probs}
+\end{figure}
+\subsection{See-Saw Neutrinos}
 The See-Saw method is the most common way of explaining the possible existence of sterile neutrinos.
@@
 This means that we have the normal neutrino eigenstate and a sterile neutrino eigenstate.
+One can see that as $M$ increases or decreases $m_D^2/M$ does the opposite -- thus the name ``see-saw.'' See Figure~\ref{fig:seesaw}.
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=.84\linewidth]{../figures/seesaw.pdf}
+  \caption{An illustration of the see-saw mechanism with left handed and right handed neutrino masses}
+  \label{fig:seesaw}
+\end{figure}
+
+There are different see-saw models associated with the introduction of isospin singlets, scalar triplets, and fermion triplets for types I, II, and III respectively.
+The Type I See-Saw model is by far the most popular with theorists.
+It suffers from requiring an incredibly large majorana mass around $10^{13}$ GeV making it untestable in current particle accelerators~\cite{justin2}.
+Type II models also rely on quite large Majorana masses.
+Type III models have the best chance at experimental verification but are the least phenomenologically satisfying and are thus the least popular.
 
@@
 \section{Conclusions}
-xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx xxxxxxxx xxxxxx
+The discovery of heavy, right-handed sterile neutrinos would help explain some of the mystery surrounding the nature of neutrinos.
+While there is some evidence for sterile neutrinos from LSND and MiniBooNE, it is very hotly contested evidence.
+Further work remains to be done before the answer is clear.
+Neutrino specific experiments like MINOS+, Super-K, and Daya Bay are continuing the search for evidence for sterile neutrinos.
+As the amount of data continues to increase, the anomalies will certainly be confirmed or refuted very soon.
 
@@
   T2K Experiment -- \url{http://t2k-experiment.org/neutrinos/in-the-standard-model/}
+\bibitem{ping2}
+Solar neutrinos. Rep. Prog. Phys. 64 (2001) 97–146
+\bibitem{ping3}
+Daya Bay Experiment: \url{https://www.bnl.gov/science/dayabay.php}
+\bibitem{ping4}
+Neutrinoless double beta decay and CP violation Physics Letters B Volume 157, Issues 5–6, 25 July 1985, Pages 442–446
+\bibitem{ping5}
+Neutrinoless Double Beta Decay Experiments. arXiv:1408.2455v1 [hep-ex] 11 Aug 2014
 
@@
 \bibitem{justin1}
-Biswajoy Brahmachari, A see-saw model of sterile neutrino, \emph{Physics Letters B}, {\bf 461} 3, 26 August 1999, Pages 243-247.
+Biswajoy Brahmachari, A see-saw model of sterile neutrino, \emph{Phys. Lett. B,}, {\bf 461} 3, 26 August 1999, Pages 243-247.
+\bibitem{justin2}
+W. Grimus, L. Lavoura, B. Radovcic, Type II seesaw mechanism for Higgs doublets and the scale of new physics, \emph{Phys. Lett. B.}, {\bf 674}, 2, 13 April 2009, Pages 117-121.
 \bibitem{LSND}

### 2015-04-15 p3_505/presentation/pres.tex
typos

@@
     \end{align}
-  \item And therefore a complete expression without hte complex exponential is:
+  \item And therefore a complete expression without the complex exponential is:
     \begin{align}
@@
         \item Cherenkov detector using 1220 PMTs.
-        \item Protons strike target, use pions decaying at test to product $\overline{\nu}_\mu$.
+        \item Protons strike target, use pions decaying at rest to product $\overline{\nu}_\mu$.
         \end{itemize}
@@
    \end{itemize}
-  \item NOvA: Long baseline experiment using the NuMI beam
+  \item NO$\nu$A: Long baseline experiment using the NuMI beam
    \begin{itemize}
@@
   \item Some, very contested, evidence for sterile neutrinos has been found at LSND and MiniBooNE
-  \item MINOS+, Super-K, Daya Bay, MicroBooNE, and NoVA will continue searching for sterile neutrinos
+  \item MINOS+, Super-K, Daya Bay, MicroBooNE, and NO$\nu$A will continue searching for sterile neutrinos
   \item Expect to see the anomalies be confirmed or refuted in the next year or two
